---
title: "TMAO and longitudinal cognitive function"
author: "P. Lee"
date: "September 2026 revision"
output:
  html_document:
    toc: true
    toc_depth: 4
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE, cache = FALSE)

library(dplyr)
library(readr)
library(nlme)
library(knitr)
library(ggplot2)
library(ggfortify)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(knitr::kable(out, align = c("l", "r", "r", "c", "r", "r", "r")))
}

model_n_by_quintile <- function(model, data, id = "pat.id", quintile = "tmao_q",
                                extra_vars = character()) {
  vars <- unique(c(all.vars(formula(model)), id, quintile, extra_vars))
  vars <- vars[vars %in% names(data)]

  used <- data %>%
    filter(if_all(all_of(vars), ~ !is.na(.x)))

  used %>%
    group_by(.data[[quintile]]) %>%
    summarise(
      N_patients = n_distinct(.data[[id]]),
      N_observations = n(),
      .groups = "drop"
    )
}

print_model_set <- function(models) {
  for (nm in names(models)) {
    cat("\n\n#### ", nm, "\n\n", sep = "")
    print_lme_fixed(models[[nm]])
  }
}

print_model_n_set <- function(models, data, extra_vars = character()) {
  out <- bind_rows(lapply(names(models), function(nm) {
    model_n_by_quintile(models[[nm]], data = data, extra_vars = extra_vars) %>%
      mutate(Model = nm, .before = 1)
  }))
  print(knitr::kable(out))
}

```

# Data preparation

```{r data-preparation, results='asis'}

baseline_raw <- read_csv("Baseline.csv", show_col_types = FALSE)

baseline_tmao <- baseline_raw %>%
  transmute(
    pat.id = as.character(pat.id),
    tmao_q = as.factor(tmao_q)
  ) %>%
  distinct()

if (anyDuplicated(baseline_tmao$pat.id)) {
  stop("Baseline.csv contains more than one TMAO-quintile value for at least one pat.id.")
}

attach_baseline_tmao_q <- function(data, baseline_lookup = baseline_tmao) {
  data %>%
    mutate(pat.id = as.character(pat.id)) %>%
    select(-any_of("tmao_q")) %>%
    left_join(baseline_lookup, by = "pat.id")
}

check_tmao_q_fixed <- function(data, dataset_name = deparse(substitute(data))) {
  changing <- data %>%
    filter(!is.na(tmao_q)) %>%
    group_by(pat.id) %>%
    summarise(n_tmao_q = n_distinct(tmao_q), .groups = "drop") %>%
    filter(n_tmao_q > 1)

  if (nrow(changing) > 0) {
    stop(dataset_name, ": tmao_q is not fixed within participant.")
  }
  invisible(TRUE)
}

baseline_tmao %>%
  count(tmao_q, name = "N") %>%
  knitr::kable()

factor_vars <- c(
  "pat.id", "visit.name", "pat.sex", "vhf.typ.aktuell.bl", "rauchen",
  "stop.grund", "prev.hypertonie", "prev.diabetes", "prev.stroke.tia",
  "prev.coronary.heart.disease", "sport", "highest.education.level.groups",
  "type_of_lesion", "dementia", "cog_impairment", "MoCA_change",
  "CoCo_change", "DSST_change", "SF_change", "TMTA_change", "TMTB_change"
)

factorise_analysis_vars <- function(data) {
  data %>% mutate(across(any_of(factor_vars), as.factor))
}

fill_baseline_egfr <- function(data) {
  data %>%
    group_by(pat.id) %>%
    mutate(
      baseline_creatinine = dplyr::first(
        creatinine.gfr[as.character(visit.name) == "Baseline"],
        default = NA_real_
      ),
      creatinine.gfr = ifelse(
        is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr
      )
    ) %>%
    ungroup() %>%
    select(-baseline_creatinine)
}


dd <- read_csv("Fu5.csv", show_col_types = FALSE) %>%
  attach_baseline_tmao_q() %>%
  mutate(tmao_log = log1p(tmao)) %>%
  factorise_analysis_vars() %>%
  fill_baseline_egfr()

check_tmao_q_fixed(dd, "Fu5.csv")
```

# Model definitions

The six adjustment models retain the original Models 1–3 and incorporate the September 2026 metabolic updates:

- **Model 1:** age, sex, education, quality of life, GDS.
- **Model 2:** Model 1 + smoking, BMI, hypertension, physical activity, alcohol.
- **Model 3:** Model 2 + coronary artery disease, previous stroke/TIA.
- **Model 4a:** Model 3 + eGFR.
- **Model 4b:** Model 3 + diabetes.
- **Model 5:** Model 3 + diabetes + eGFR.

```{r model-specification}

cog_covariates <- list(
  "Model 1" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total"
  ),
  "Model 2" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption"
  ),
  "Model 3" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia"
  ),
  "Model 4a" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia", "creatinine.gfr"
  ),
  "Model 4b" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia", "prev.diabetes"
  ),
  "Model 5" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia",
    "prev.diabetes", "creatinine.gfr"
  )
)

build_longitudinal_formula <- function(outcome, exposure, covariates) {
  rhs <- c(
    "time_since_first",
    exposure,
    paste0("time_since_first:", exposure),
    covariates
  )
  as.formula(paste(outcome, "~", paste(rhs, collapse = " + ")))
}

fit_cognitive_lme_set <- function(outcome, exposure, data,
                                  random = ~ time_since_first | center/pat.id,
                                  covariates = cog_covariates) {
  setNames(lapply(names(covariates), function(nm) {
    nlme::lme(
      fixed = build_longitudinal_formula(outcome, exposure, covariates[[nm]]),
      random = random,
      data = data,
      na.action = na.exclude
    )
  }), names(covariates))
}

```

# Association between TMAO and overall cognitive function over time

All primary models are linear mixed-effects models with a random intercept and random slope for time nested within centre/patient, as in the supplied longitudinal analysis.

## CoCo score

### TMAO quintiles

```{r coco-quintile, results='asis'}
coco_q <- fit_cognitive_lme_set("CoCo", "tmao_q", dd)
print_model_set(coco_q)

cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(coco_q, dd, extra_vars = c("center", "pat.id"))
```

### TMAO continuous

```{r coco-continuous, results='asis'}
coco_cont <- fit_cognitive_lme_set("CoCo", "tmao_log", dd)
print_model_set(coco_cont)

cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(coco_cont, dd, extra_vars = c("center", "pat.id"))
```

## MoCA score

### TMAO quintiles

```{r moca-quintile, results='asis'}
moca_q <- fit_cognitive_lme_set("MoCA", "tmao_q", dd)
print_model_set(moca_q)

cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(moca_q, dd, extra_vars = c("center", "pat.id"))
```

### TMAO continuous

```{r moca-continuous, results='asis'}
moca_cont <- fit_cognitive_lme_set("MoCA", "tmao_log", dd)
print_model_set(moca_cont)

cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(moca_cont, dd, extra_vars = c("center", "pat.id"))
```

### One diagnostic plot for MoCA

`autoplot()` is applied to a fixed-effects-only `lm` companion model.

```{r moca-autoplot, fig.width=10, fig.height=8}
moca_diag_formula <- build_longitudinal_formula(
  "MoCA", "tmao_q", cog_covariates[["Model 5"]]
)

moca_diag_lm <- lm(
  formula = moca_diag_formula,
  data = dd,
  na.action = na.exclude
)

autoplot(moca_diag_lm, which = c(1, 2, 3, 5), ncol = 2)
```

## TMT-A score

### TMAO quintiles

```{r tmta-quintile, results='asis'}
tmta_q <- fit_cognitive_lme_set("TMTA", "tmao_q", dd)
print_model_set(tmta_q)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(tmta_q, dd, extra_vars = c("center", "pat.id"))
```

### TMAO continuous

```{r tmta-continuous, results='asis'}
tmta_cont <- fit_cognitive_lme_set("TMTA", "tmao_log", dd)
print_model_set(tmta_cont)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(tmta_cont, dd, extra_vars = c("center", "pat.id"))
```

## TMT-B score

### TMAO quintiles

```{r tmtb-quintile, results='asis'}
tmtb_q <- fit_cognitive_lme_set("TMTB", "tmao_q", dd)
print_model_set(tmtb_q)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(tmtb_q, dd, extra_vars = c("center", "pat.id"))
```

### TMAO continuous

```{r tmtb-continuous, results='asis'}
tmtb_cont <- fit_cognitive_lme_set("TMTB", "tmao_log", dd)
print_model_set(tmtb_cont)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(tmtb_cont, dd, extra_vars = c("center", "pat.id"))
```

## Semantic fluency score

### TMAO quintiles

```{r sf-quintile, results='asis'}
sf_q <- fit_cognitive_lme_set("SF", "tmao_q", dd)
print_model_set(sf_q)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(sf_q, dd, extra_vars = c("center", "pat.id"))
```

### TMAO continuous

```{r sf-continuous, results='asis'}
sf_cont <- fit_cognitive_lme_set("SF", "tmao_log", dd)
print_model_set(sf_cont)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(sf_cont, dd, extra_vars = c("center", "pat.id"))
```

## DSST score

### TMAO quintiles

```{r dsst-quintile, results='asis'}
dsst_q <- fit_cognitive_lme_set("DSST", "tmao_q", dd)
print_model_set(dsst_q)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(dsst_q, dd, extra_vars = c("center", "pat.id"))
```

### TMAO continuous

```{r dsst-continuous, results='asis'}
dsst_cont <- fit_cognitive_lme_set("DSST", "tmao_log", dd)
print_model_set(dsst_cont)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(dsst_cont, dd, extra_vars = c("center", "pat.id"))
```
