---
title: "TMAO and cognitive decline: survival, sensitivity, and stroke-subgroup analyses"
author: "P. Lee"
date: "September 2026 revision"
output:
  html_document:
    toc: true
    toc_depth: 4
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE, cache = FALSE)

library(dplyr)
library(readr)
library(nlme)
library(survival)
library(knitr)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(knitr::kable(out, align = c("l", "r", "r", "c", "r", "r", "r")))
}

model_n_by_quintile <- function(model, data, id = "pat.id", quintile = "tmao_q",
                                extra_vars = character()) {
  vars <- unique(c(all.vars(formula(model)), id, quintile, extra_vars))
  vars <- vars[vars %in% names(data)]

  used <- data %>%
    filter(if_all(all_of(vars), ~ !is.na(.x)))

  used %>%
    group_by(.data[[quintile]]) %>%
    summarise(
      N_patients = n_distinct(.data[[id]]),
      N_observations = n(),
      .groups = "drop"
    )
}

print_model_set <- function(models) {
  for (nm in names(models)) {
    cat("\n\n#### ", nm, "\n\n", sep = "")
    print_lme_fixed(models[[nm]])
  }
}

print_model_n_set <- function(models, data, extra_vars = character()) {
  out <- bind_rows(lapply(names(models), function(nm) {
    model_n_by_quintile(models[[nm]], data = data, extra_vars = extra_vars) %>%
      mutate(Model = nm, .before = 1)
  }))
  print(knitr::kable(out))
}

# Numeric fixed-effect table for survival::coxph models.
print_cox_table <- function(model, digits = 3) {
  beta <- stats::coef(model)
  vc <- stats::vcov(model)
  se <- sqrt(diag(vc))
  z <- beta / se
  p <- 2 * stats::pnorm(abs(z), lower.tail = FALSE)
  lower <- beta - stats::qnorm(0.975) * se
  upper <- beta + stats::qnorm(0.975) * se

  fmt_num <- function(x) formatC(x, format = "f", digits = digits)
  fmt_p <- function(x) ifelse(
    is.na(x), NA_character_,
    ifelse(x < 0.001, "<0.001", formatC(x, format = "f", digits = digits))
  )

  out <- data.frame(
    Parameter = names(beta),
    HR = fmt_num(exp(beta)),
    `95% CI` = paste0(fmt_num(exp(lower)), " to ", fmt_num(exp(upper))),
    `p value` = fmt_p(p),
    check.names = FALSE,
    row.names = NULL
  )

  print(knitr::kable(out, align = c("l", "r", "c", "r")))
}

print_cox_set <- function(models) {
  for (nm in names(models)) {
    cat("\n\n#### ", nm, "\n\n", sep = "")
    print_cox_table(models[[nm]])
  }
}

```

# Data preparation

```{r data-preparation, results='asis'}

# September 2026 correction:
# TMAO quintile is a BASELINE exposure. It is never recalculated in longitudinal,
# follow-up, subgroup, or sensitivity-analysis datasets.
baseline_raw <- read_csv("Baseline.csv", show_col_types = FALSE)

baseline_tmao <- baseline_raw %>%
  transmute(
    pat.id = as.character(pat.id),
    tmao_q = as.factor(tmao_q)
  ) %>%
  distinct()

if (anyDuplicated(baseline_tmao$pat.id)) {
  stop("Baseline.csv contains more than one TMAO-quintile value for at least one pat.id.")
}

attach_baseline_tmao_q <- function(data, baseline_lookup = baseline_tmao) {
  data %>%
    mutate(pat.id = as.character(pat.id)) %>%
    select(-any_of("tmao_q")) %>%
    left_join(baseline_lookup, by = "pat.id")
}

check_tmao_q_fixed <- function(data, dataset_name = deparse(substitute(data))) {
  changing <- data %>%
    filter(!is.na(tmao_q)) %>%
    group_by(pat.id) %>%
    summarise(n_tmao_q = n_distinct(tmao_q), .groups = "drop") %>%
    filter(n_tmao_q > 1)

  if (nrow(changing) > 0) {
    stop(dataset_name, ": tmao_q is not fixed within participant.")
  }
  invisible(TRUE)
}

# Baseline quintile counts used in the manuscript cohort.
baseline_tmao %>%
  count(tmao_q, name = "N") %>%
  knitr::kable()

factor_vars <- c(
  "pat.id", "visit.name", "pat.sex", "vhf.typ.aktuell.bl", "rauchen",
  "stop.grund", "prev.hypertonie", "prev.diabetes", "prev.stroke.tia",
  "prev.coronary.heart.disease", "sport", "highest.education.level.groups",
  "type_of_lesion", "dementia", "cog_impairment", "MoCA_change",
  "CoCo_change", "DSST_change", "SF_change", "TMTA_change", "TMTB_change"
)

factorise_analysis_vars <- function(data) {
  data %>% mutate(across(any_of(factor_vars), as.factor))
}

# As in the September 2026 analysis, carry the baseline eGFR value to rows
# where longitudinal creatinine.gfr is missing.
fill_baseline_egfr <- function(data) {
  data %>%
    group_by(pat.id) %>%
    mutate(
      baseline_creatinine = dplyr::first(
        creatinine.gfr[as.character(visit.name) == "Baseline"],
        default = NA_real_
      ),
      creatinine.gfr = ifelse(
        is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr
      )
    ) %>%
    ungroup() %>%
    select(-baseline_creatinine)
}


dd <- read_csv("Fu5.csv", show_col_types = FALSE) %>%
  attach_baseline_tmao_q() %>%
  mutate(tmao_log = log1p(tmao)) %>%
  factorise_analysis_vars() %>%
  fill_baseline_egfr()

FUP3 <- read_csv("FUP3.long.csv", show_col_types = FALSE) %>%
  attach_baseline_tmao_q() %>%
  mutate(tmao_log = log1p(tmao)) %>%
  factorise_analysis_vars() %>%
  fill_baseline_egfr()

WGS <- read_csv("WGS.long.csv", show_col_types = FALSE) %>%
  attach_baseline_tmao_q() %>%
  mutate(tmao_log = log1p(tmao)) %>%
  factorise_analysis_vars() %>%
  fill_baseline_egfr()

WGS$time_to_CoCo_change <- as.numeric(WGS$time_to_CoCo_change)

check_tmao_q_fixed(dd, "Fu5.csv")
check_tmao_q_fixed(FUP3, "FUP3.long.csv")
check_tmao_q_fixed(WGS, "WGS.long.csv")
```

# Model definitions

```{r model-specification}

cog_covariates <- list(
  "Model 1" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total"
  ),
  "Model 2" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption"
  ),
  "Model 3" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia"
  ),
  "Model 4a" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia", "creatinine.gfr"
  ),
  "Model 4b" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia", "prev.diabetes"
  ),
  "Model 5" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia",
    "prev.diabetes", "creatinine.gfr"
  )
)

build_longitudinal_formula <- function(outcome, exposure, covariates) {
  rhs <- c(
    "time_since_first",
    exposure,
    paste0("time_since_first:", exposure),
    covariates
  )
  as.formula(paste(outcome, "~", paste(rhs, collapse = " + ")))
}

fit_cognitive_lme_set <- function(outcome, exposure, data,
                                  random = ~ time_since_first | center/pat.id,
                                  covariates = cog_covariates) {
  setNames(lapply(names(covariates), function(nm) {
    nlme::lme(
      fixed = build_longitudinal_formula(outcome, exposure, covariates[[nm]]),
      random = random,
      data = data,
      na.action = na.exclude
    )
  }), names(covariates))
}


build_cox_formula <- function(time_var, event_var, exposure, covariates) {
  lhs <- paste0(
    "Surv(", time_var, ", as.numeric(as.character(", event_var, ")))"
  )
  rhs <- c(exposure, covariates, "frailty(center)")
  as.formula(paste(lhs, "~", paste(rhs, collapse = " + ")))
}

fit_cox_set <- function(time_var, event_var, exposure, data,
                        covariates = cog_covariates) {
  setNames(lapply(names(covariates), function(nm) {
    survival::coxph(
      formula = build_cox_formula(
        time_var, event_var, exposure, covariates[[nm]]
      ),
      data = data,
      na.action = na.exclude,
      model = TRUE
    )
  }), names(covariates))
}

print_cox_n_set <- function(models, data) {
  out <- bind_rows(lapply(names(models), function(nm) {
    model_n_by_quintile(
      models[[nm]], data = data,
      extra_vars = c("center", "pat.id")
    ) %>% mutate(Model = nm, .before = 1)
  }))
  print(knitr::kable(out))
}
```

# Association between TMAO and time to cognitive decline

The same six covariate sets are used for all time-to-event outcomes.

```{r event-runner, include=FALSE}
run_event_analysis <- function(title, time_var, event_var, data = dd) {
  cat("\n\n## ", title, "\n\n", sep = "")

  cat("\n\n### TMAO quintiles\n\n")
  q_models <- fit_cox_set(time_var, event_var, "tmao_q", data)
  print_cox_set(q_models)
  cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
  print_cox_n_set(q_models, data)

  cat("\n\n### TMAO continuous\n\n")
  c_models <- fit_cox_set(time_var, event_var, "tmao_log", data)
  print_cox_set(c_models)
  cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
  print_cox_n_set(c_models, data)

  invisible(list(quintile = q_models, continuous = c_models))
}
```

```{r main-events, results='asis'}
event_mci <- run_event_analysis(
  "Mild cognitive impairment (MoCA < 26)",
  "time_to_cog_impairment", "cog_impairment", dd
)

event_dementia <- run_event_analysis(
  "Early dementia (MoCA < 21)",
  "time_to_dementia", "dementia", dd
)

event_coco <- run_event_analysis(
  "CoCo score decline >= 1 SD",
  "time_to_CoCo_change", "CoCo_change", dd
)

event_tmta <- run_event_analysis(
  "TMT-A decline >= 1 SD",
  "time_to_TMTA_change", "TMTA_change", dd
)

event_tmtb <- run_event_analysis(
  "TMT-B decline >= 1 SD",
  "time_to_TMTB_change", "TMTB_change", dd
)

event_sf <- run_event_analysis(
  "Semantic fluency decline >= 1 SD",
  "time_to_SF_change", "SF_change", dd
)

event_dsst <- run_event_analysis(
  "DSST decline >= 1 SD",
  "time_to_DSST_change", "DSST_change", dd
)

event_moca <- run_event_analysis(
  "MoCA decline >= 1 SD",
  "time_to_MoCA_change", "MoCA_change", dd
)
```

# Sensitivity analysis: long-term and short-term groups

## Long-term group (FUP3)

### Longitudinal CoCo and MoCA

```{r fup3-lme, results='asis'}
cat("\n\n#### CoCo: TMAO quintiles\n\n")
fup3_coco_q <- fit_cognitive_lme_set("CoCo", "tmao_q", FUP3)
print_model_set(fup3_coco_q)
print_model_n_set(fup3_coco_q, FUP3, extra_vars = c("center", "pat.id"))

cat("\n\n#### CoCo: TMAO continuous\n\n")
fup3_coco_cont <- fit_cognitive_lme_set("CoCo", "tmao_log", FUP3)
print_model_set(fup3_coco_cont)
print_model_n_set(fup3_coco_cont, FUP3, extra_vars = c("center", "pat.id"))

cat("\n\n#### MoCA: TMAO quintiles\n\n")
fup3_moca_q <- fit_cognitive_lme_set("MoCA", "tmao_q", FUP3)
print_model_set(fup3_moca_q)
print_model_n_set(fup3_moca_q, FUP3, extra_vars = c("center", "pat.id"))

cat("\n\n#### MoCA: TMAO continuous\n\n")
fup3_moca_cont <- fit_cognitive_lme_set("MoCA", "tmao_log", FUP3)
print_model_set(fup3_moca_cont)
print_model_n_set(fup3_moca_cont, FUP3, extra_vars = c("center", "pat.id"))
```

### Long-term time-to-event analyses

```{r fup3-events, results='asis'}
fup3_mci <- run_event_analysis(
  "FUP3: Mild cognitive impairment (MoCA < 26)",
  "time_to_cog_impairment", "cog_impairment", FUP3
)

fup3_coco_event <- run_event_analysis(
  "FUP3: CoCo score decline >= 1 SD",
  "time_to_CoCo_change", "CoCo_change", FUP3
)
```

## Short-term group (WGS)


```{r wgs-events, results='asis'}
wgs_mci <- run_event_analysis(
  "WGS: Mild cognitive impairment (MoCA < 26)",
  "time_to_cog_impairment", "cog_impairment", WGS
)
```

# Stroke subgroup analyses

The supplied stroke-subgroup scripts omit previous stroke/TIA from Model 3 onward because stroke phenotype defines the subgroup. The random structure for the subgroup longitudinal MoCA models remains a random intercept within centre/patient, matching the supplied subgroup file.

```{r stroke-model-specification}
cog_stroke_covariates <- list(
  "Model 1" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total"
  ),
  "Model 2" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption"
  ),
  "Model 3" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease"
  ),
  "Model 4a" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "creatinine.gfr"
  ),
  "Model 4b" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.diabetes"
  ),
  "Model 5" = c(
    "age.bl", "pat.sex", "highest.education.level.groups",
    "gesundheitszustand", "gds.total",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.diabetes", "creatinine.gfr"
  )
)

ddns <- dd %>% filter(type_of_lesion == "No Stroke")
ddos <- dd %>% filter(type_of_lesion == "Overt Stroke")
ddss <- dd %>% filter(type_of_lesion == "Silent Stroke")
```

## Longitudinal MoCA according to stroke subtype

```{r stroke-moca, results='asis'}
for (subgroup_name in c("No Stroke", "Overt Stroke", "Silent Stroke")) {
  subgroup_data <- switch(
    subgroup_name,
    "No Stroke" = ddns,
    "Overt Stroke" = ddos,
    "Silent Stroke" = ddss
  )

  cat("\n\n### ", subgroup_name, "\n\n", sep = "")

  cat("\n\n#### TMAO quintiles\n\n")
  q_models <- fit_cognitive_lme_set(
    "MoCA", "tmao_q", subgroup_data,
    random = ~ 1 | center/pat.id,
    covariates = cog_stroke_covariates
  )
  print_model_set(q_models)
  print_model_n_set(q_models, subgroup_data, extra_vars = c("center", "pat.id"))

  cat("\n\n#### TMAO continuous\n\n")
  c_models <- fit_cognitive_lme_set(
    "MoCA", "tmao_log", subgroup_data,
    random = ~ 1 | center/pat.id,
    covariates = cog_stroke_covariates
  )
  print_model_set(c_models)
  print_model_n_set(c_models, subgroup_data, extra_vars = c("center", "pat.id"))
}
```

## Mild cognitive impairment according to stroke subtype

```{r stroke-mci, results='asis'}
for (subgroup_name in c("No Stroke", "Overt Stroke", "Silent Stroke")) {
  subgroup_data <- switch(
    subgroup_name,
    "No Stroke" = ddns,
    "Overt Stroke" = ddos,
    "Silent Stroke" = ddss
  )

  cat("\n\n### ", subgroup_name, "\n\n", sep = "")

  q_models <- fit_cox_set(
    "time_to_cog_impairment", "cog_impairment", "tmao_q",
    subgroup_data, covariates = cog_stroke_covariates
  )
  cat("\n\n#### TMAO quintiles\n\n")
  print_cox_set(q_models)
  print_cox_n_set(q_models, subgroup_data)

  c_models <- fit_cox_set(
    "time_to_cog_impairment", "cog_impairment", "tmao_log",
    subgroup_data, covariates = cog_stroke_covariates
  )
  cat("\n\n#### TMAO continuous\n\n")
  print_cox_set(c_models)
  print_cox_n_set(c_models, subgroup_data)
}
```

## MoCA decline >= 1 SD according to stroke subtype

The supplied stroke-subgroup file contains this analysis for **No Stroke** and **Overt Stroke**

```{r stroke-moca-decline, results='asis'}
for (subgroup_name in c("No Stroke", "Overt Stroke")) {
  subgroup_data <- switch(
    subgroup_name,
    "No Stroke" = ddns,
    "Overt Stroke" = ddos
  )

  cat("\n\n### ", subgroup_name, "\n\n", sep = "")

  q_models <- fit_cox_set(
    "time_to_MoCA_change", "MoCA_change", "tmao_q",
    subgroup_data, covariates = cog_stroke_covariates
  )
  cat("\n\n#### TMAO quintiles\n\n")
  print_cox_set(q_models)
  print_cox_n_set(q_models, subgroup_data)

  c_models <- fit_cox_set(
    "time_to_MoCA_change", "MoCA_change", "tmao_log",
    subgroup_data, covariates = cog_stroke_covariates
  )
  cat("\n\n#### TMAO continuous\n\n")
  print_cox_set(c_models)
  print_cox_n_set(c_models, subgroup_data)
}
```
