---
title: "TMAO and hippocampal volume"
author: "P. Lee"
date: "September 2026 revision"
output:
  html_document:
    toc: true
    toc_depth: 4
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE, cache = FALSE)

library(dplyr)
library(readr)
library(nlme)
library(knitr)
library(ggplot2)
library(ggfortify)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(knitr::kable(out, align = c("l", "r", "r", "c", "r", "r", "r")))
}

model_n_by_quintile <- function(model, data, id = "pat.id", quintile = "tmao_q",
                                extra_vars = character()) {
  vars <- unique(c(all.vars(formula(model)), id, quintile, extra_vars))
  vars <- vars[vars %in% names(data)]

  used <- data %>%
    filter(if_all(all_of(vars), ~ !is.na(.x)))

  used %>%
    group_by(.data[[quintile]]) %>%
    summarise(
      N = n_distinct(.data[[id]]),
      .groups = "drop"
    )
}

print_model_set <- function(models) {
  for (nm in names(models)) {
    cat("\n\n#### ", nm, "\n\n", sep = "")
    print_lme_fixed(models[[nm]])
  }
}

print_model_n_set <- function(models, data, extra_vars = character()) {
  out <- bind_rows(lapply(names(models), function(nm) {
    model_n_by_quintile(
      models[[nm]],
      data = data,
      extra_vars = extra_vars
    ) %>%
      mutate(Model = nm, .before = 1)
  }))

  print(knitr::kable(
    out,
    col.names = c("Model", "TMAO quintile", "N")
  ))
}

```

# Data preparation

```{r data-preparation, results='asis'}

baseline_raw <- read_csv("Baseline.csv", show_col_types = FALSE)

baseline_tmao <- baseline_raw %>%
  transmute(
    pat.id = as.character(pat.id),
    tmao_q = as.factor(tmao_q)
  ) %>%
  distinct()

if (anyDuplicated(baseline_tmao$pat.id)) {
  stop("Baseline.csv contains more than one TMAO-quintile value for at least one pat.id.")
}

attach_baseline_tmao_q <- function(data, baseline_lookup = baseline_tmao) {
  data %>%
    mutate(pat.id = as.character(pat.id)) %>%
    select(-any_of("tmao_q")) %>%
    left_join(baseline_lookup, by = "pat.id")
}

check_tmao_q_fixed <- function(data, dataset_name = deparse(substitute(data))) {
  changing <- data %>%
    filter(!is.na(tmao_q)) %>%
    group_by(pat.id) %>%
    summarise(n_tmao_q = n_distinct(tmao_q), .groups = "drop") %>%
    filter(n_tmao_q > 1)

  if (nrow(changing) > 0) {
    stop(dataset_name, ": tmao_q is not fixed within participant.")
  }
  invisible(TRUE)
}

# Baseline quintile counts used in the manuscript cohort.
baseline_tmao %>%
  count(tmao_q, name = "N") %>%
  knitr::kable()


dd <- baseline_raw %>%
  mutate(
    pat.id = as.character(pat.id),
    tmao_q = as.factor(tmao_q),
    tmao_log = log1p(tmao),
    type_of_lesion = as.factor(type_of_lesion)
  )

# Follow-up/sensitivity datasets inherit the TMAO quintile from Baseline.csv.
fup3 <- read_csv("FUP3.long.csv", show_col_types = FALSE) %>%
  attach_baseline_tmao_q() %>%
  mutate(tmao_log = log1p(tmao))

wgs <- read_csv("WGS.long.csv", show_col_types = FALSE) %>%
  attach_baseline_tmao_q() %>%
  mutate(tmao_log = log1p(tmao))

check_tmao_q_fixed(dd, "Baseline.csv")
check_tmao_q_fixed(fup3, "FUP3.csv")
check_tmao_q_fixed(wgs, "WGS.csv")
```

# Model definitions

```{r model-specification}
hv_covariates <- list(
  "Model 1" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups"
  ),
  "Model 2" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption"
  ),
  "Model 3" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia"
  ),
  "Model 4a" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia", "creatinine.gfr"
  ),
  "Model 4b" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia", "prev.diabetes"
  ),
  "Model 5" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.stroke.tia",
    "prev.diabetes", "creatinine.gfr"
  )
)

build_cross_sectional_formula <- function(outcome, exposure, covariates) {
  as.formula(
    paste(outcome, "~", paste(c(exposure, covariates), collapse = " + "))
  )
}

fit_hv_set <- function(outcome, exposure, data,
                       covariates = hv_covariates) {
  setNames(lapply(names(covariates), function(nm) {
    nlme::lme(
      fixed = build_cross_sectional_formula(
        outcome, exposure, covariates[[nm]]
      ),
      random = ~ 1 | center,
      data = data,
      na.action = na.exclude
    )
  }), names(covariates))
}
```

# Association between TMAO and hippocampal volume

Explanatory variables are baseline TMAO quintiles and continuous log-transformed TMAO. Random intercepts are specified for centre.

## Bilateral hippocampal volume

### TMAO quintiles

```{r hvraw-quintile, results='asis'}
hvraw_q <- fit_hv_set("HVraw", "tmao_q", dd)
print_model_set(hvraw_q)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(hvraw_q, dd, extra_vars = "center")
```

### TMAO continuous

```{r hvraw-continuous, results='asis'}
hvraw_cont <- fit_hv_set("HVraw", "tmao_log", dd)
print_model_set(hvraw_cont)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(hvraw_cont, dd, extra_vars = "center")
```

### One diagnostic plot for bilateral hippocampal volume

```{r hvraw-autoplot, fig.width=10, fig.height=8}
hv_diag_lm <- lm(
  formula = build_cross_sectional_formula(
    "HVraw", "tmao_q", hv_covariates[["Model 5"]]
  ),
  data = dd,
  na.action = na.exclude
)

ggfortify::autoplot(hv_diag_lm, which = c(1, 2, 3, 5), ncol = 2)
```

## Left hippocampal volume

### TMAO quintiles

```{r lhv-quintile, results='asis'}
lhv_q <- fit_hv_set("LHV", "tmao_q", dd)
print_model_set(lhv_q)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(lhv_q, dd, extra_vars = "center")
```

### TMAO continuous

```{r lhv-continuous, results='asis'}
lhv_cont <- fit_hv_set("LHV", "tmao_log", dd)
print_model_set(lhv_cont)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(lhv_cont, dd, extra_vars = "center")
```

## Right hippocampal volume

### TMAO quintiles

```{r rhv-quintile, results='asis'}
rhv_q <- fit_hv_set("RHV", "tmao_q", dd)
print_model_set(rhv_q)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(rhv_q, dd, extra_vars = "center")
```

### TMAO continuous

```{r rhv-continuous, results='asis'}
rhv_cont <- fit_hv_set("RHV", "tmao_log", dd)
print_model_set(rhv_cont)
cat("\n\n#### Analysis N by baseline TMAO quintile\n\n")
print_model_n_set(rhv_cont, dd, extra_vars = "center")
```

# Stroke subgroup analyses

As in the supplied hippocampal-volume subgroup code, previous stroke/TIA is not included as an adjustment variable within stroke-defined subgroups.

```{r stroke-model-specification}
hv_stroke_covariates <- list(
  "Model 1" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups"
  ),
  "Model 2" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption"
  ),
  "Model 3" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease"
  ),
  "Model 4a" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "creatinine.gfr"
  ),
  "Model 4b" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.diabetes"
  ),
  "Model 5" = c(
    "age.bl", "pat.sex", "sbtiv.vol", "highest.education.level.groups",
    "rauchen", "bmi", "prev.hypertonie", "sport", "total.alcohol.consumption",
    "prev.coronary.heart.disease", "prev.diabetes", "creatinine.gfr"
  )
)

dos <- dd %>% filter(type_of_lesion == "Overt Stroke")
dss <- dd %>% filter(type_of_lesion == "Silent Stroke")
dns <- dd %>% filter(type_of_lesion == "No Stroke")
```

```{r hv-stroke-subgroups, results='asis'}
for (subgroup_name in c("Overt Stroke", "Silent Stroke", "No Stroke")) {
  subgroup_data <- switch(
    subgroup_name,
    "Overt Stroke" = dos,
    "Silent Stroke" = dss,
    "No Stroke" = dns
  )

  cat("\n\n## ", subgroup_name, "\n\n", sep = "")

  cat("\n\n### TMAO quintiles\n\n")
  q_models <- fit_hv_set(
    "HVraw", "tmao_q", subgroup_data,
    covariates = hv_stroke_covariates
  )
  print_model_set(q_models)
  print_model_n_set(q_models, subgroup_data, extra_vars = "center")

  cat("\n\n### TMAO continuous\n\n")
  c_models <- fit_hv_set(
    "HVraw", "tmao_log", subgroup_data,
    covariates = hv_stroke_covariates
  )
  print_model_set(c_models)
  print_model_n_set(c_models, subgroup_data, extra_vars = "center")
}
```

# Sensitivity analyses: long-term and short-term groups

The same six-model sequence is applied to the long-term (`FUP3.csv`) and short-term (`WGS.csv`) hippocampal-volume sensitivity datasets. These datasets now inherit their quintile from `Baseline.csv` rather than recalculating quintiles within the smaller follow-up samples.

```{r hv-sensitivity, results='asis'}
for (analysis_name in c("Long-term group", "Short-term group")) {
  analysis_data <- switch(
    analysis_name,
    "Long-term group" = fup3,
    "Short-term group" = wgs
  )

  cat("\n\n## ", analysis_name, "\n\n", sep = "")

  cat("\n\n### TMAO quintiles\n\n")
  q_models <- fit_hv_set("HVraw", "tmao_q", analysis_data)
  print_model_set(q_models)
  print_model_n_set(q_models, analysis_data, extra_vars = "center")

  cat("\n\n### TMAO continuous\n\n")
  c_models <- fit_hv_set("HVraw", "tmao_log", analysis_data)
  print_model_set(c_models)
  print_model_n_set(c_models, analysis_data, extra_vars = "center")
}
```
