---
title: "Stroke subgroup analyses"
author: "P. Lee"
date: "11.03.2025"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE
)

library(tidyverse)
library(nlme)
library(survival)
library(knitr)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
# Uses nlme directly to avoid version-sensitive extraction of random-effect
# variances by parameters/datawizard. Sample size is deliberately not printed.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "r", "c", "r", "r", "r")
    )
  )
}

# Numeric fixed-effect table for survival::coxph models.
# Avoids forestmodel/ggplot compatibility issues and deliberately does not
# print sample size or event counts.
print_cox_table <- function(model, digits = 3) {
  beta <- stats::coef(model)
  vc <- stats::vcov(model)

  common <- intersect(names(beta), rownames(vc))
  beta <- beta[common]
  se <- sqrt(diag(vc[common, common, drop = FALSE]))

  z <- beta / se
  p <- 2 * stats::pnorm(abs(z), lower.tail = FALSE)
  lower <- exp(beta - stats::qnorm(0.975) * se)
  upper <- exp(beta + stats::qnorm(0.975) * se)

  fmt_num <- function(x) {
    ifelse(
      is.na(x),
      NA_character_,
      formatC(x, format = "f", digits = digits)
    )
  }

  fmt_p <- function(x) {
    ifelse(
      is.na(x),
      NA_character_,
      ifelse(
        x < 0.001,
        "<0.001",
        formatC(x, format = "f", digits = digits)
      )
    )
  }

  out <- data.frame(
    Parameter = names(beta),
    HR = fmt_num(exp(beta)),
    `95% CI` = paste0(fmt_num(lower), " to ", fmt_num(upper)),
    `p value` = fmt_p(p),
    check.names = FALSE,
    row.names = NULL
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "c", "r")
    )
  )
}
```

```{r data-wrangling}
dd <- read_csv("Fu5.csv")
dd$tmao_log <- log1p(dd$tmao )
dd$tmao_q <- as.factor(dd$tmao_q)
dd$center <- as.factor(dd$center)

dd <- dd %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

ddns <- dd %>%
  filter(type_of_lesion == 'No Stroke')
ddos <- dd %>%
  filter(type_of_lesion == 'Overt Stroke')
ddss <- dd %>%
  filter(type_of_lesion == 'Silent Stroke')
```

# Mixed effect models of MoCA score and TMAO according to stroke subtype

## No Stroke

### TMAO Quintiles

* Model 1
```{r m1, results='asis'}
m1 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand+
                 gds.total,
          random = ~ 1 | center/pat.id, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m1)
```

* Model 2
```{r m2, results='asis'}
m2 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ 1 | center/pat.id, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m2)
```

* Model 3
```{r m3, results='asis'}
m3 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease,
          random = ~ 1 | center/pat.id, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m3)
```

\newpage

### TMAO Continuous

* Model 1
```{r m10, results='asis'}
m10 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ 1 | center/pat.id, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m10)
```

* Model 2
```{r m20, results='asis'}
m20 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ 1 | center/pat.id, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m20)
```

* Model 3
```{r m30, results='asis'}
m30 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease ,
          random = ~ 1 | center/pat.id, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m30)
```

\newpage

## Overt Stroke

### TMAO Quintiles

* Model 1
```{r m5, results='asis'}
m5 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand+
                 gds.total,
          random = ~ 1 | center/pat.id, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m5)
```

* Model 2
```{r m6, results='asis'}
m6 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ 1 | center/pat.id, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m6)
```

* Model 3
```{r m7, results='asis'}
m7 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease ,
          random = ~ 1 | center/pat.id, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m7)
```


\newpage

### TMAO Continuous

* Model 1
```{r m50, results='asis'}
m50 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ 1 | center/pat.id, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m50)
```

* Model 2
```{r m60, results='asis'}
m60 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ 1 | center/pat.id, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m60)
```

* Model 3
```{r m70, results='asis'}
m70 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease ,
          random = ~ 1 | center/pat.id, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m70)
```

\newpage

## Silent Stroke

### TMAO Quintiles

* Model 1
```{r m15, results='asis'}
m15 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand+
                 gds.total,
          random = ~ 1 | center/pat.id, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m15)
```

* Model 2
```{r m16, results='asis'}
m16 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ 1 | center/pat.id, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m16)
```

* Model 3
```{r m17, results='asis'}
m17 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease,
          random = ~ 1 | center/pat.id, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m17)
```



\newpage

### TMAO Continuous

* Model 1
```{r m150, results='asis'}
m150 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ 1 | center/pat.id, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m150)
```

* Model 2
```{r m160, results='asis'}
m160 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ 1 | center/pat.id, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m160)
```

* Model 3
```{r m170, results='asis'}
m170 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease,
          random = ~ 1 | center/pat.id, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m170)
```


\newpage

# Cox proportional hazards regression model

## Mild cognitive impairment (MoCA < 26)

### No Stroke

#### TMAO Quintiles

```{r c1234,results="asis", fig.width=10, fig.height=8,fig.align='center', out.width='100%'}

c1 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddns)

print_cox_table(c1)

c2 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddns) 

print_cox_table(c2)

c3 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease 
             + frailty(center), data = ddns) 

print_cox_table(c3)

c4 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = ddns) 

print_cox_table(c4)

```

\newpage

#### TMAO continuous

```{r c12340,results="asis", fig.width=10, fig.height=8, fig.align='center', out.width='100%'}

c10 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddns)

print_cox_table(c10)

c20 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddns) 

print_cox_table(c20)

c30 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddns) 

print_cox_table(c30)

c40 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = ddns) 

print_cox_table(c40)

```

\newpage

### Overt Stroke

#### TMAO Quintiles

```{r c5678,results="asis", fig.width=10, fig.height=8, fig.align='center', out.width='100%'}

c5 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddos)

print_cox_table(c5)

c6 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddos) 

print_cox_table(c6)

c7 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddos) 

print_cox_table(c7)

c8 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = ddos) 

print_cox_table(c8)

```

\newpage

#### TMAO continuous

```{r c56780,results="asis", fig.width=10, fig.height=8, fig.align='center', out.width='100%'}

c50 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddos)

print_cox_table(c50)

c60 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddos) 

print_cox_table(c60)

c70 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddos) 

print_cox_table(c70)

c80 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = ddos) 

print_cox_table(c80)

```

\newpage

### Silent Stroke

#### TMAO Quintiles

```{r c15161718,results="asis", fig.width=10, fig.height=8, fig.align='center', out.width='100%'}

c15 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddss)

print_cox_table(c15)

c16 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddss) 

print_cox_table(c16)

c17 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddss) 

print_cox_table(c17)

c18 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = ddss) 

print_cox_table(c18)

```

\newpage

#### TMAO continuous

```{r c151617180,results="asis", fig.width=10, fig.height=8, fig.align='center', out.width='100%'}

c150 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddss)

print_cox_table(c150)

c160 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddss) 

print_cox_table(c160)

c170 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddss) 

print_cox_table(c170)

c180 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = ddss) 

print_cox_table(c180)

```

\newpage

## Cognitive change (MoCA change >= 1 SD)

### No Stroke

#### TMAO Quintiles

```{r c35363738,results="asis", fig.width=10, fig.height=8, fig.align='center', out.width='100%'}

c35 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddns)

print_cox_table(c35)

c36 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddns) 

print_cox_table(c36)

c37 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddns) 

print_cox_table(c37)

```

\newpage

#### TMAO continuous
```{r c353637380,results="asis", fig.width=12, fig.height=8, fig.align='center', out.width='100%'}

c350 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddns)

print_cox_table(c350)

c360 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddns) 

print_cox_table(c360)

c370 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddns) 

print_cox_table(c370)

```

\newpage

### Overt Stroke

#### TMAO Quintiles

```{r c25262728,results="asis", fig.width=12, fig.height=8, fig.align='center', out.width='100%'}

c25 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddos)

print_cox_table(c25)

c26 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddos) 

print_cox_table(c26)

c27 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddos) 

print_cox_table(c27)

```

\newpage

#### TMAO continuous
```{r c252627280,results="asis", fig.width=12, fig.height=8, fig.align='center', out.width='100%'}

c250 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = ddos)

print_cox_table(c250)

c260 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = ddos) 

print_cox_table(c260)

c270 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease   
             + frailty(center), data = ddos) 

print_cox_table(c270)

```

\newpage
