---
title: "Association between Trimethylamine N-Oxide (TMAO) and longitudinal cognitive function  in Patients with Atrial Fibrillation - Part 1: Longitudinal LME analyses"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
  html_document:
    toc: true
    toc_depth: '4'
    df_print: paged
header-includes: \usepackage[some]{background}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, error=F, message=F, warning=F, cache=T)

library(dplyr)
library(readr)
library(nlme)
library(knitr)
options(scipen=999)


# Fixed-effect table for nlme::lme models.
# Uses nlme directly and deliberately does not print sample size.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    tt$`p-value` < 0.001,
    "<0.001",
    formatC(tt$`p-value`, format = "f", digits = digits)
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "r", "c", "r", "r", "r")
    )
  )
}
```

```{r Data Wrangling}
dd <- read_csv("Fu5.csv")
dd$tmao_log <- log(dd$tmao + 1)
dd$tmao_q <- as.factor(dd$tmao_q)

dd <- dd %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

dd <- dd %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)
```

\newpage

# Association between TMAO and overall cognitive function over time (LME)

**linear mixed effects model (LME)** is used according to manual page 6 (structure model I)  

* Covariates
Model 1 (epidemiologic variables): age, sex, education (years), quality of life (EQ-ED-5L), geriatric depression scale (GDS)  
Model 2: (modifiable cardiovascular risk factors): Model 1 + smoking status, BMI, hypertension, alcohol consumption, physical activity  
Model 3 (cardiovascular events): Model 1 + Model 2 + coronary artery disease, prev. stroke  
Model 4a (metabolic variables): Model 1 + Model 2 + Model 3 + renal function (eGFR)  
Model 4b (metabolic variables): Model 1 + Model 2 + Model 3 + diabetes status  
Model 5 (metabolic variables): Model 1 + Model 2 + Model 3 + diabetes status, renal function (eGFR)  


\newpage

## CoCo score

### TMAO quintiles

* Model 4a
```{r m4a, results='asis'}
m4a <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4a)
```


* Model 4b
```{r m4b, results='asis'}
m4b <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4b)
```


\newpage

### TMAO Continuous

* Model 1
```{r m10, results='asis'}
m10 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m10) 
```

* Model 2
```{r m20, results='asis'}
m20 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m20)
```

* Model 3
```{r m30, results='asis'}
m30 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m30) 
```

* Model 4a
```{r m4a0, results='asis'}
m4a0 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4a0)
```


* Model 4b
```{r m4b0, results='asis'}
m4b0 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes ,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4b0)
```

* Model 5
```{r m40, results='asis'}
m40 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m40)
```

\newpage

## MoCA score

### TMAO Quintiles

* Model 4a
```{r m8a, results='asis'}
m8a <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m8a)
```

* Model 4b
```{r m8b, results='asis'}
m8b <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m8b)
```

\newpage

### TMAO Continuous

* Model 1
```{r m50, results='asis'}
m50 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m50)
```

* Model 2
```{r m60, results='asis'}
m60 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m60)
```

* Model 3
```{r m70, results='asis'}
m70 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m70) 
```


* Model 4a
```{r m8a0, results='asis'}
m8a0 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m8a0)
```

* Model 4b
```{r m8b0, results='asis'}
m8b0 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m8b0)
```

* Model 5
```{r m80, results='asis'}
m80 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m80)
```

\newpage

## TMT-A score

### TMAO Quintiles 

* Model 4a
```{r m14a, results='asis'}
m14a <- lme(TMTA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m14a)
```

* Model 4b
```{r m14b, results='asis'}
m14b <- lme(TMTA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m14b)
```

\newpage

### TMAO Continuous

* Model 1
```{r m110, results='asis'}
m110 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m110)
```

* Model 2
```{r m120, results='asis'}
m120 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m120) 
```

* Model 3
```{r m130, results='asis'}
m130 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m130) 
```


* Model 4a
```{r m14a0, results='asis'}
m14a0 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m14a0)
```

* Model 4b
```{r m14b0, results='asis'}
m14b0 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m14b0)
```

* Model 5
```{r m140, results='asis'}
m140 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m140)
```

\newpage

## TMT-B score

### TMAO Quintiles


* Model 4a
```{r m18a, results='asis'}
m18a <- lme(TMTB ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m18a)
```

* Model 4b
```{r m18b, results='asis'}
m18b <- lme(TMTB ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m18b)
```

\newpage

### TMAO Continuous

* Model 1
```{r m150, results='asis'}
m150 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m150) 
```

* Model 2
```{r m160, results='asis'}
m160 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m160) 
```

* Model 3
```{r m170, results='asis'}
m170 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m170) 
```


* Model 4a
```{r m18a0, results='asis'}
m18a0 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m18a0)
```

* Model 4b
```{r m18b0, results='asis'}
m18b0 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m18b0)
```

* Model 5
```{r m180, results='asis'}
m180 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m180) 
```

\newpage

## SF score

### TMAO Quintiles

* Model 4a
```{r m24a, results='asis'}
m24a <- lme(SF ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m24a)
```

* Model 4b
```{r m24b, results='asis'}
m24b <- lme(SF ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m24b)
```

\newpage

### TMAO Continuous

* Model 1
```{r m210, results='asis'}
m210 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m210) 
```

* Model 2
```{r m220, results='asis'}
m220 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m220)
```

* Model 3
```{r m230, results='asis'}
m230 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m230) 
```

* Model 4a
```{r m24a0, results='asis'}
m24a0 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m24a0)
```

* Model 4b
```{r m24b0, results='asis'}
m24b0 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m24b0)
```

* Model 5
```{r m240, results='asis'}
m240 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m240) 
```


\newpage 

## DSST score

### TMAO Quintiles

* Model 4a
```{r m28a, results='asis'}
m28a <- lme(DSST ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m28a)
```

* Model 4b
```{r m28b, results='asis'}
m28b <- lme(DSST ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m28b)
```

\newpage

### TMAO Continuous

* Model 1
```{r m250, results='asis'}
m250 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m250) 
```

* Model 2
```{r m260, results='asis'}
m260 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m260) 
```

* Model 3
```{r m270, results='asis'}
m270 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m270) 
```

* Model 4a
```{r m28a0, results='asis'}
m28a0 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m28a0)
```

* Model 4b
```{r m28b0, results='asis'}
m28b0 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m28b0)
```

* Model 5
```{r m280, results='asis'}
m280 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m280) 
```
