---
title: "Association between Trimethylamine N-Oxide (TMAO) and Neurofilament light polypeptide in Patients with Atrial Fibrillation"
author: "P. Lee"
date: "18.03.2025"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
  html_document:
    toc: true
    toc_depth: 4
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE
)

library(tidyverse)
library(nlme)
library(knitr)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
# Uses nlme directly to avoid version-sensitive extraction of random-effect
# variances by parameters/datawizard. Sample size is deliberately not printed.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "r", "c", "r", "r", "r")
    )
  )
}
```

```{r data-wrangling}
dd <- read_csv("Baseline.csv")

dd <- dd %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
dd$tmao_log <- log1p(dd$tmao )
dd$tmao_q <- as.factor(dd$tmao_q)

```

\newpage

# TMAO Quintiles and Neurofilament light polypeptide

```{r table, results='asis'}
ddq <- dd %>%
  pivot_longer(cols=c(tmao_q),
               names_to = "TMAO",values_to = "Quintiles")%>%
  pivot_longer(cols = c(nfl),
               names_to = "nfl",values_to = "Volumes")

### Rearrange order of factor 
ddq$Quintiles <- factor(ddq$Quintiles, 
                       levels = c("1", "2", "3", "4", '5'))

ddq <- ddq %>%
  filter(!is.na(Quintiles)) %>%
  filter(!is.na(Volumes))

sddq <- ddq %>%
  group_by(Quintiles, nfl, TMAO) %>%
  summarise(n = n(), mean = mean(Volumes), sd = sd(Volumes), .groups = "drop")

kable(sddq)

```
```{r boxplot, fig.show='asis', fig.width=8, fig.height=8}
a<- ggplot(ddq, aes(x=Quintiles,y=Volumes,fill=Quintiles))+
  geom_boxplot()+
  xlab("") +
  ylab("") +
  ggtitle("Neurofilament light polypeptide") +
  scale_fill_brewer(palette="YlOrRd") +
  theme_bw()+
  theme(legend.position = "none") +
  theme(strip.text = element_text(color = "black")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
a

```

\newpage

# Association between TMAO and Neurofilament light polypeptide (Nfl)

* Random Effects  
Center  

* Covariates  
Model 1 (epidemiologic variables): age, sex  
Model 2: (modifiable cardiovascular risk factors): Model 1 + BMI, hypertension  
Model 3 (cardiovascular events): Model 1 + Model 2 + coronary artery disease, prev. stroke   
Model 4 (metabolic variables): Model 1 + Model 2 + diabetes status, renal function (eGFR)  

\newpage

## TMAO quintiles

* Model 1
```{r m1, results='asis'}
m1 <- lme(nfl ~ tmao_q + age.bl + pat.sex,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m1)
```

* Model 2
```{r m2, results='asis'}
m2 <- lme(nfl ~ tmao_q + age.bl + pat.sex +
                 bmi + prev.hypertonie,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m2)
```

* Model 3
```{r m3, results='asis'}
m3 <- lme(nfl ~ tmao_q + age.bl + pat.sex +
                 bmi + prev.hypertonie
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m3)
```

* Model 4
```{r m4, results='asis'}
m4 <- lme(nfl ~ tmao_q + age.bl + pat.sex +
                 bmi + prev.hypertonie
                 + prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4)
```
\newpage

## TMAO Continuous

* Model 1
```{r m10, results='asis'}
m10 <- lme(nfl ~ tmao_log + age.bl + pat.sex,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m10)
```

* Model 2
```{r m20, results='asis'}
m20 <- lme(nfl ~ tmao_log + age.bl + pat.sex +
                 bmi + prev.hypertonie,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m20)
```

* Model 3
```{r m30, results='asis'}
m30 <- lme(nfl ~ tmao_log + age.bl + pat.sex +
                 bmi + prev.hypertonie
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m30)
```

* Model 4
```{r m40, results='asis'}
m40 <- lme(nfl ~ tmao_log + age.bl + pat.sex +
                 bmi + prev.hypertonie
                 + prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m40)
```