---
title: "Association between Trimethylamine N-Oxide (TMAO) and Hippocampal Volume  in Patients with Atrial Fibrillation"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
  html_document:
    toc: true
    toc_depth: '4'
    df_print: paged
header-includes: \usepackage[some]{background}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, error=F, message=F, warning=F, cache=T)

library(tidyverse)
library(GGally)
library(nlme)
library(forestplot)
library(forestploter)
library(ggridges)
library(viridis)
library(viridisLite)
library(hrbrthemes)
library(wesanderson)
library(ggfortify)
library(ggthemes)
library(RColorBrewer)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(lme4)
library(cowplot)
library(ggpubr)
library(knitr)
library(arsenal)
library(ggcorrplot)
library(psych)
library(PerformanceAnalytics)
library(easystats)
library(RColorBrewer)
options(scipen=999)
```

```{r Data Wrangling}
dd <- read_csv("Baseline.csv")

dd <- dd %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
dd$tmao_log <- log(dd$tmao + 1)
dd$tmao_q <- as.factor(dd$tmao_q)

fup3 <- read_csv("FUP3.csv")
fup3 <- fup3 %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
fup3$tmao_log <- log(fup3$tmao + 1)
fup3$tmao_q <- as.factor(fup3$tmao_q)

wgs <- read_csv("WGS.csv")
wgs <- wgs %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
wgs$tmao_log <- log(wgs$tmao + 1)
wgs$tmao_q <- as.factor(wgs$tmao_q)

```

\newpage


# Association between TMAO and hippocampal volume

Explanatory variables: TMAO quintiles, TMAO continuous (log-transform)  
Response variables: HVraw, LHV, RHV, AI  
Random intercepts: Center  


Model 4a (Metabolic variables): Model 3 + eGFR + Diabetes status
Model 4b (Metabolic variables): Model 3 + Diabetes status
Model 5 (Metabolic variables): Model 3 + eGFR + Diabetes status  

## Bilateral hippocampal volumes

### TMAO Quintiles

**Table 3: Association of bilateral hippocampal volumes and tmao quintiles**


* Model 4a
```{r m4a, results='asis'}
m4a <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4a, effects = "fixed") |> print_md()
```


* Model 4b
```{r m4b, results='asis'}
m4b <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4b, effects = "fixed") |> print_md()
```


* Model 5
```{r m5, results='asis'}
m5 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m5, effects = "fixed") |> print_md()
```
\newpage

### TMAO Continuous

**Table 4: Association of bilateral hippocampal volumes and tmao continuous (log-transform)**

* Model 4a
```{r m4a0, results='asis'}
m4a0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4a0, effects = "fixed") |> print_md()
```

* Model 4b
```{r m4b0, results='asis'}
m4b0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4b0, effects = "fixed") |> print_md()
```

* Model 5
```{r m50, results='asis'}
m50 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m50, effects = "fixed") |> print_md()
```

\newpage

## Left hippocampal volumes

### TMAO Quintiles

**Table 5: Association of left hippocampal volumes and TMAO quintiles**


* Model 4a
```{r m4al, results='asis'}
m4al <- lme(LHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4al, effects = "fixed") |> print_md()
```

* Model 4b
```{r m4bl, results='asis'}
m4bl <- lme(LHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4bl, effects = "fixed") |> print_md()
```

* Model 5
```{r m5l, results='asis'}
m5l <- lme(LHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m5l, effects = "fixed") |> print_md()
```

\newpage

### TMAO Continuous

**Table 6: Association of left hippocampal volumes and tmao continuous (log-transform)**


* Model 4a
```{r m4a0l, results='asis'}
m4a0l <- lme(LHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4a0l, effects = "fixed") |> print_md()
```


* Model 4b
```{r m4b0l, results='asis'}
m4b0l <- lme(LHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4b0l, effects = "fixed") |> print_md()
```

* Model 5
```{r m50l, results='asis'}
m50l <- lme(LHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m50l, effects = "fixed") |> print_md()
```

\newpage

## Right hippocampal volumes

### TMAO Quintiles 

**Table 7: Association of right hippocampal volumes and TMAO quintiles**

* Model 4a
```{r m4ar, results='asis'}
m4ar <- lme(RHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4ar, effects = "fixed") |> print_md()
```

* Model 4b
```{r m4br, results='asis'}
m4br <- lme(RHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4br, effects = "fixed") |> print_md()
```

* Model 5
```{r m5r, results='asis'}
m5r <- lme(RHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m5r, effects = "fixed") |> print_md()
```

\newpage

### TMAO Continuous

**Table 8: Association of right hippocampal volumes and tmao continuous (log-transform)**

* Model 4a
```{r m4a0r, results='asis'}
m4a0r <- lme(RHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4a0r, effects = "fixed") |> print_md()
```

* Model 4b
```{r m4b0r, results='asis'}
m4b0r <- lme(RHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m4b0r, effects = "fixed") |> print_md()
```

* Model 5
```{r m50r, results='asis'}
m50r <- lme(RHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

model_parameters(m50r, effects = "fixed") |> print_md()
```

\newpage


\newpage 

# Stroke subgroup analyses

Explanatory variables: TMAO quintiles, TMAO continuous (log-transform)    
Response variables: HVraw  
Random intercepts: Center   

## Overt stroke (n=480)

### TMAO Quintiles

* Model 4a
```{r m9a, results='asis'}
dd$type_of_lesion <- as.factor(dd$type_of_lesion)
dos <- dd %>%
  filter(type_of_lesion == 'Overt Stroke')

m9a <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
         + creatinine.gfr,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

model_parameters(m9a, effects = "fixed") |> print_md()
```

* Model 4b
```{r m9b, results='asis'}
dd$type_of_lesion <- as.factor(dd$type_of_lesion)
dos <- dd %>%
  filter(type_of_lesion == 'Overt Stroke')

m9b <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

model_parameters(m9b, effects = "fixed") |> print_md()
```

* Model 5
```{r m95, results='asis'}
dd$type_of_lesion <- as.factor(dd$type_of_lesion)
dos <- dd %>%
  filter(type_of_lesion == 'Overt Stroke')

m95 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

model_parameters(m95, effects = "fixed") |> print_md()
```

\newpage

### TMAO Continuous

* Model 4a
```{r m90a, results='asis'}
m90a <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          + creatinine.gfr,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

model_parameters(m90a, effects = "fixed") |> print_md()
```

* Model 4b
```{r m90b, results='asis'}
m90b <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

model_parameters(m90b, effects = "fixed") |> print_md()
```

* Model 5
```{r m950, results='asis'}
m950 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

model_parameters(m950, effects = "fixed") |> print_md()
```
\newpage

## Silent stroke (n=399)

### TMAO Quintiles

* Model 4a
```{r m14a, results='asis'}
dss <- dd %>%
  filter(type_of_lesion == 'Silent Stroke')
m14a <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          + creatinine.gfr,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

model_parameters(m14a, effects = "fixed") |> print_md()
```

* Model 4b
```{r m14b, results='asis'}
dss <- dd %>%
  filter(type_of_lesion == 'Silent Stroke')
m14b <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

model_parameters(m14b, effects = "fixed") |> print_md()
```

* Model 5
```{r m145, results='asis'}
m145 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

model_parameters(m145, effects = "fixed") |> print_md()
```

\newpage

### TMAO Continuous

* Model 4a
```{r m14a0, results='asis'}
m14a0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
         + creatinine.gfr,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

model_parameters(m14a0, effects = "fixed") |> print_md()
```

* Model 4b
```{r m14b0, results='asis'}
m14b0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

model_parameters(m14b0, effects = "fixed") |> print_md()
```

* Model 5
```{r m1450, results='asis'}
m1450 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

model_parameters(m1450, effects = "fixed") |> print_md()
```


\newpage

## No stroke (n=994)

### TMAO Quintiles

* Model 4a
```{r m18a, results='asis'}
dns <- dd %>%
  filter(type_of_lesion == 'No Stroke')
m18a <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          + creatinine.gfr,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

model_parameters(m18a, effects = "fixed") |> print_md()
```

* Model 4b
```{r m18b, results='asis'}
m18b <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

model_parameters(m18b, effects = "fixed") |> print_md()
```

* Model 5
```{r m185, results='asis'}
m185 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

model_parameters(m185, effects = "fixed") |> print_md()
```

\newpage

### TMAO Continuous


* Model 4a
```{r m18a0, results='asis'}
m18a0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          + creatinine.gfr,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

model_parameters(m18a0, effects = "fixed") |> print_md()
```

* Model 4b
```{r m18b0, results='asis'}
m18b0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

model_parameters(m18b0, effects = "fixed") |> print_md()
```


* Model 5
```{r m1805, results='asis'}
m1805 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

model_parameters(m1805, effects = "fixed") |> print_md()
```


\newpage


# Sensitivity analyses for TMAO with HV for LTG and STG  

Model 4a: model 3 + eGFR  
Model 4b: model 3 + DM 

## Long-term group  
### TMAO Quintiles  

* Model 4a
```{r m24a, results='asis'}
m24a <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          + creatinine.gfr,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

model_parameters(m24a, effects = "fixed") |> print_md()
```

* Model 4b
```{r m24b, results='asis'}
m24b <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes ,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

model_parameters(m24b, effects = "fixed") |> print_md()
```
\newpage

### TMAO Continuous  

* Model 4a
```{r m24a0, results='asis'}
m24a0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          + creatinine.gfr,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

model_parameters(m24a0, effects = "fixed") |> print_md()
```

* Model 4b
```{r m24b0, results='asis'}
m24b0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes ,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

model_parameters(m24b0, effects = "fixed") |> print_md()
```

\newpage

## Short term group  

### TMAO Quintiles

* Model 4a
```{r m28a, results='asis'}
m28a <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
           + creatinine.gfr,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

model_parameters(m28a, effects = "fixed") |> print_md()
```


* Model 4b
```{r m28b, results='asis'}
m28b <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes ,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

model_parameters(m28b, effects = "fixed") |> print_md()
```

\newpage

### TMAO Continuous

* Model 4a
```{r m28a0, results='asis'}
m28a0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
           + creatinine.gfr,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

model_parameters(m28a0, effects = "fixed") |> print_md()
```

* Model 4b
```{r m28b0, results='asis'}
m28b0 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

model_parameters(m28b0, effects = "fixed") |> print_md()
```

\newpage