---
title: "Association between Trimethylamine N-Oxide (TMAO) and Neurofilament light polypeptide in Patients with Atrial Fibrillation"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
  html_document:
    toc: true
    toc_depth: '4'
    df_print: paged
header-includes: \usepackage[some]{background}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, error=F, message=F, warning=F, cache=T)

library(readr)
library(dplyr)
library(nlme)
library(knitr)

options(scipen = 999)

# Robust fixed-effect table for nlme::lme models.
# This deliberately avoids parameters::model_parameters(), because some
# parameters/datawizard versions fail while trying to extract random-effect
# variances from lme objects ("more names in select than in replacement").
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  # Coefficient-specific 95% Wald CI using the denominator df from nlme.
  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    tt$`p-value` < 0.001,
    "<0.001",
    formatC(tt$`p-value`, format = "f", digits = digits)
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(knitr::kable(out, align = c("l", "r", "r", "c", "r", "r", "r")))
}
```

```{r Data Wrangling}
dd <- read_csv("Baseline.csv")

dd <- dd %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
dd$tmao_log <- log(dd$tmao + 1)
dd$tmao_q <- as.factor(dd$tmao_q)

```

\newpage


# Association between TMAO and Neurofilament light polypeptide (Nfl)

* Random Effects  
Center  

* Covariates  
Model 1 (epidemiologic variables): age, sex  
Model 2: (modifiable cardiovascular risk factors): Model 1 + BMI, hypertension  
Model 3 (cardiovascular events): Model 1 + Model 2 + coronary artery disease, prev. stroke  
Model 4a (metabolic variables): Model 1 + Model 2 + renal function (eGFR)  
Model 4b (metabolic variables): Model 1 + Model 2 + diabetes status  
Model 5 (metabolic variables): Model 1 + Model 2 + diabetes status, renal function (eGFR)  

\newpage

## TMAO quintiles

* Model 4a
```{r m4a, results='asis'}
m4a <- lme(nfl ~ tmao_q + age.bl + pat.sex +
                 bmi + prev.hypertonie
           + prev.coronary.heart.disease + prev.stroke.tia
                  + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4a)
```

* Model 4b
```{r m4b, results='asis'}
m4b <- lme(nfl ~ tmao_q + age.bl + pat.sex +
                 bmi + prev.hypertonie
           + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4b)
```

* Model 5
```{r m5, results='asis'}
m5 <- lme(nfl ~ tmao_q + age.bl + pat.sex +
                 bmi + prev.hypertonie
           + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m5)
```
\newpage

## TMAO Continuous

* Model 4a
```{r m4a0, results='asis'}
m4a0 <- lme(nfl ~ tmao_log + age.bl + pat.sex +
                 bmi + prev.hypertonie
            + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4a0)
```

* Model 4b
```{r m4b0, results='asis'}
m4b0 <- lme(nfl ~ tmao_log + age.bl + pat.sex +
                 bmi + prev.hypertonie
            + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4b0)
```

* Model 5
```{r m50, results='asis'}
m50 <- lme(nfl ~ tmao_log + age.bl + pat.sex +
                 bmi + prev.hypertonie
            + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m50)
```
