---
title: "Association between Trimethylamine N-Oxide (TMAO) (Quintiles) and Hippocampal Volume  in Patients with Atrial Fibrillation"
author: "P. Lee"
date: "28.2.2025"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
  html_document:
    toc: true
    toc_depth: 4
    df_print: paged
header-includes: \usepackage[some]{background}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE
)

library(tidyverse)
library(nlme)
library(ggpubr)
library(knitr)
library(arsenal)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
# This avoids parameters::model_parameters(), which can fail with some
# parameters/datawizard versions when extracting random-effect variances.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "r", "c", "r", "r", "r")
    )
  )
}
```

```{r data-wrangling}
dd <- read_csv("Baseline.csv") %>%
  mutate(
    tmao_q = factor(ntile(tmao, 5)),
    tmao_log = log1p(tmao)
  )

FUP3 <- read_csv("FUP3.csv") %>%
  mutate(
    tmao_q = factor(ntile(tmao, 5)),
    tmao_log = log1p(tmao)
  )

WGS <- read_csv("WGS.csv") %>%
  mutate(
    tmao_q = factor(ntile(tmao, 5)),
    tmao_log = log1p(tmao)
  )
```

\newpage

# Baseline characteristics according to TMAO Quintiles

**Table 1: Baseline characteristics of patients according to TMAO quintiles**

```{r baseline, results="asis"}
tab1 <- tableby(
  tmao_q ~ nfl + HVraw + LHV + RHV + AI + sbtiv.vol +
    lncci.presence + snci.presence + mb.presence + wml.presence +
    wml.vol + mod.faz + brain.volume.raw + brain.volume.normalized +
    CoCo + MoCA + SF + DSST + TMTA + TMTB +
    duration + stop.grund + vhf.typ.aktuell.bl + age.bl + pat.sex +
    highest.education.level.groups + gesundheitszustand +
    gds.total + rauchen + bmi + prev.hypertonie +
    total.alcohol.consumption + essen.fruechte + essen.gemuese +
    essen.fleisch + essen.milch.portion + sport +
    prev.stroke.tia + prev.coronary.heart.disease + prev.diabetes +
    creatinine.gfr + ldl + hdl + trigly + chol + chads +
    med.oak.name + med.antiplatelet.yn,
  data = dd
)

kable(summary(tab1))
```

\newpage

## TMAO Quintiles and Hippocampal volumes

**Table 2: TMAO quintiles and Hippocampal volumes**

```{r table-hv, results='asis'}
ddq <- dd %>%
  pivot_longer(cols=c(tmao_q),
               names_to = "TMAO",values_to = "Quintiles")%>%
  pivot_longer(cols = c(HVraw, AI, LHV, RHV, sbtiv.vol),
               names_to = "HV",values_to = "Volumes")

### Rearrange order of factor 
ddq$Quintiles <- factor(ddq$Quintiles, 
                       levels = c("1", "2", "3", "4", '5'))

ddq <- ddq %>%
  filter(!is.na(Quintiles)) %>%
  filter(!is.na(Volumes))

sddq <- ddq %>%
  group_by(Quintiles, HV, TMAO) %>%
  summarise(n = n(), mean = mean(Volumes), sd = sd(Volumes), .groups = "drop")

kable(sddq)

```
unit = mm^3^  

HVraw: Bilateral hippocampal volume (LHV + RHV)  
AI: Asymmetry index  
LHV: Left hippocampal volume  
RHV: Right hippocampal volume  
sbtiv.vol: Intracranial volume  

\newpage

**Figure 1: Boxplot of hippocampal volume according to TMAO quintiles**

```{r boxplot, fig.show='asis', fig.width=12, fig.height=5}
hvraw_plot <- ddq %>%
  filter(HV == "HVraw")

plot_hvraw <- ggplot(
  hvraw_plot,
  aes(x = Quintiles, y = Volumes, fill = Quintiles)
) +
  geom_boxplot() +
  labs(x = NULL, y = NULL, title = "Bilateral Hippocampal Volume") +
  scale_fill_brewer(palette = "YlOrRd") +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

lhv_plot <- ddq %>%
  filter(HV == "LHV")

plot_lhv <- ggplot(
  lhv_plot,
  aes(x = Quintiles, y = Volumes, fill = Quintiles)
) +
  geom_boxplot() +
  labs(x = NULL, y = NULL, title = "Left Hippocampal Volume") +
  scale_fill_brewer(palette = "YlOrRd") +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

rhv_plot <- ddq %>%
  filter(HV == "RHV")

plot_rhv <- ggplot(
  rhv_plot,
  aes(x = Quintiles, y = Volumes, fill = Quintiles)
) +
  geom_boxplot() +
  labs(x = NULL, y = NULL, title = "Right Hippocampal Volume") +
  scale_fill_brewer(palette = "YlOrRd") +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggarrange(
  plot_hvraw,
  plot_lhv,
  plot_rhv,
  labels = c("A", "B", "C"),
  ncol = 3,
  nrow = 1
)
```

**Figure 2: Boxplot of intracranial volume according to TMAO quintiles**

```{r boxplot-icv, fig.show='asis', fig.width=5, fig.height=5}
icv_plot <- ddq %>%
  filter(HV == "sbtiv.vol")

ggplot(
  icv_plot,
  aes(x = Quintiles, y = Volumes, fill = Quintiles)
) +
  geom_boxplot() +
  labs(x = NULL, y = NULL, title = "Intracranial Volume") +
  scale_fill_brewer(palette = "YlOrRd") +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

\newpage

# Association between TMAO and hippocampal volume

Explanatory variables: TMAO quintiles, TMAO continuous (log-transform)  
Response variables: HVraw, LHV, RHV, AI  
Random intercepts: Center  

Model 1 (Epidemiologic variables): Age, Sex, Level of education, Intracranial Volume (sbtiv.vol)  
Model 2 (Modifiable cardiovascular risk factors): Model 1 + smoking, BMI, HT, Alcohol consumption, Physical activity  
Model 3 (Cardiovascular events): Model 1 + Model 2 + Coronary artery disease, prev. stroke  
Model 4 (Metabolic variables): Model 1 + Model 2 + Model 3 + Diabetes status  

**I didn't include these two variables: Quality of life (EQ-ED-5L), Geriatric depression scale (GDS) in the model as it might be too many variables and the results became all non-significant**  

## Bilateral hippocampal volumes

### TMAO Quintiles

**Table 3: Association of bilateral hippocampal volumes and tmao quintiles**

* Model 1
```{r m1, results='asis'}
m1 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m1)
```

* Model 2
```{r m2, results='asis'}
m2 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m2)
```

* Model 3
```{r m3, results='asis'}
m3 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m3)
```
* Model 4
```{r m4, results='asis'}
m4 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4)
```
\newpage

### TMAO Continuous

**Table 4: Association of bilateral hippocampal volumes and tmao continuous (log-transform)**
* Model 1
```{r m10, results='asis'}
m10 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m10)
```

* Model 2
```{r m20, results='asis'}
m20 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m20)
```

* Model 3
```{r m30, results='asis'}
m30 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m30)
```

* Model 4
```{r m40, results='asis'}
m40 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m40)
```

\newpage

## Left hippocampal volumes

### TMAO Quintiles

**Table 5: Association of left hippocampal volumes and TMAO quintiles**

* Model 1
```{r m1l, results='asis'}
m1l <- lme(LHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m1l)
```

* Model 2
```{r m2l, results='asis'}
m2l <- lme(LHV ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m2l)
```

* Model 3
```{r m3l, results='asis'}
m3l <- lme(LHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m3l)
```
* Model 4
```{r m4l, results='asis'}
m4l <- lme(LHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4l)
```
\newpage

### TMAO Continuous

**Table 6: Association of left hippocampal volumes and tmao continuous (log-transform)**
* Model 1
```{r m10l, results='asis'}
m10l <- lme(LHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m10l)
```

* Model 2
```{r m20l, results='asis'}
m20l <- lme(LHV ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m20l)
```

* Model 3
```{r m30l, results='asis'}
m30l <- lme(LHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m30l)
```

* Model 4
```{r m40l, results='asis'}
m40l <- lme(LHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m40l)
```

\newpage

## Right hippocampal volumes

### TMAO Quintiles 

**Table 7: Association of right hippocampal volumes and TMAO quintiles**
* Model 1
```{r m1r, results='asis'}
m1r <- lme(RHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m1r)
```

* Model 2
```{r m2r, results='asis'}
m2r <- lme(RHV ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m2r)
```

* Model 3
```{r m3r, results='asis'}
m3r <- lme(RHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m3r)
```
* Model 4
```{r m4r, results='asis'}
m4r <- lme(RHV ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4r)
```
\newpage

### TMAO Continuous

**Table 8: Association of right hippocampal volumes and tmao continuous (log-transform)**
* Model 1
```{r m10r, results='asis'}
m10r <- lme(RHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m10r)
```

* Model 2
```{r m20r, results='asis'}
m20r <- lme(RHV~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m20r)
```

* Model 3
```{r m30r, results='asis'}
m30r <- lme(RHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m30r)
```

* Model 4
```{r m40r, results='asis'}
m40r <- lme(RHV ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m40r)
```

\newpage

## Asymmetry index

### TMAO Quintiles

**Table 9: Association of Asymmetry Index and TMAO quintiles**

* Model 1
```{r m1a, results='asis'}
m1a <- lme(AI ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m1a)
```

* Model 2
```{r m2a, results='asis'}
m2a <- lme(AI ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m2a)
```

* Model 3
```{r m3a, results='asis'}
m3a <- lme(AI ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m3a)
```
* Model 4
```{r m4a, results='asis'}
m4a <- lme(AI ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4a)
```
\newpage

### TMAO Continuous

**Table 10: Association of asymmetry index and TMAO continuous (log-transform)**
* Model 1
```{r m10a, results='asis'}
m10a <- lme(AI ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m10a)
```

* Model 2
```{r m20a, results='asis'}
m20a <- lme(AI ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m20a)
```

* Model 3
```{r m30a, results='asis'}
m30a <- lme(AI ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m30a)
```

* Model 4
```{r m40a, results='asis'}
m40a <- lme(AI ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=dd,
          na.action=na.exclude)

print_lme_fixed(m40a)
```

\newpage

# Association between total brain volume and hippocampal volume

Explanatory variables: total brain volume  
Response variables: HV raw, RHV, LHV, AI  
Random intercepts: Center    

Model 1 (Epidemiologic variables): Age, Sex, Level of education, Intracranial Volume (sbtiv.vol)   

**Bilateral hippocampal volume**

```{r m5, results='asis'}
m5 <- lme(HVraw ~ brain.volume.raw +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)
print_lme_fixed(m5)
```
**Left hippocampal volume**

```{r m5l, results='asis'}
m5l <- lme(LHV ~ brain.volume.raw +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)
print_lme_fixed(m5l)
```

**Right hippocampal volume**

```{r m5r, results='asis'}
m5r <- lme(RHV ~ brain.volume.raw +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)
print_lme_fixed(m5r)
```

**Asymmetry index**

```{r m5a, results='asis'}
m5a <- lme(AI ~ brain.volume.raw +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dd,
          na.action=na.exclude)
print_lme_fixed(m5a)
```

\newpage 

# Stroke subgroup analyses

Overt stroke: self-reported stroke or TIA event at baseline    
Silent stroke: ischemic lesions on bMRI (LNCCI and/or SNCI) + no self-reported stroke or TIA event at baseline
No stroke: no ischemic lesions on bMRI (LNCCI and/or SNCI) + no self-reported stroke or TIA event at baseline  

**Table: Stroke type and MoCA score**

```{r table-stroke, results='asis'}
dd %>%
  group_by(type_of_lesion) %>%
  summarise(
    n = n(),
    mean_moca = mean(MoCA, na.rm = TRUE)
  ) %>%
  kable(
    caption = "Number of patients and mean MoCA score by Stroke subgroup"
  )
```
Explanatory variables: TMAO quintiles, TMAO continuous (log-transform)    
Response variables: HVraw  
Random intercepts: Center   

Model 1 (Epidemiologic variables): Age, Sex, Level of education, Intracranial Volume (sbtiv.vol)    
Model 2 (Modifiable cardiovascular risk factors): Model 1 + smoking, BMI, HT, Alcohol consumption, Physical activity    
Model 3 (Cardiovascular events): Model 2 + Coronary artery disease  
Previous stroke/TIA is not added as a covariate within stroke-defined subgroups because it is part of the subgroup definition and has no useful within-group variation.  
Model 4 (Metabolic variables): Model 3 + Diabetes status  

## Overt stroke (n=480)

### TMAO Quintiles

* Model 1
```{r m6, results='asis'}
dd$type_of_lesion <- as.factor(dd$type_of_lesion)
dos <- dd %>%
  filter(type_of_lesion == 'Overt Stroke')

m6 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m6)
```

* Model 2
```{r m7, results='asis'}
m7 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m7)
```

* Model 3
```{r m8, results='asis'}
m8 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease ,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m8)
```

* Model 4
```{r m9, results='asis'}
dd$type_of_lesion <- as.factor(dd$type_of_lesion)
dos <- dd %>%
  filter(type_of_lesion == 'Overt Stroke')

m9 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m9)
```
\newpage

### TMAO Continuous

* Model 1
```{r m60, results='asis'}
m60 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m60)
```

* Model 2
```{r m70, results='asis'}
m70 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m70)
```

* Model 3
```{r m80, results='asis'}
m80 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m80)
```

* Model 4
```{r m90, results='asis'}
m90 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dos,
          na.action=na.exclude)

print_lme_fixed(m90)
```

\newpage

## Silent stroke (n=399)

### TMAO Quintiles

* Model 1
```{r m11, results='asis'}
dss <- dd %>%
  filter(type_of_lesion == 'Silent Stroke')

m11 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m11)
```

* Model 2
```{r m12, results='asis'}
m12 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m12)
```

* Model 3
```{r m13, results='asis'}
m13 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease ,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m13)
```

* Model 4
```{r m14, results='asis'}
m14 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m14)
```
\newpage

### TMAO Continuous

* Model 1
```{r m110, results='asis'}
m110 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m110)
```

* Model 2
```{r m120, results='asis'}
m120 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m120)
```

* Model 3
```{r m130, results='asis'}
m130 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease ,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m130)
```

* Model 4
```{r m140, results='asis'}
m140 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dss,
          na.action=na.exclude)

print_lme_fixed(m140)
```

\newpage

## No stroke (n=994)

### TMAO Quintiles

* Model 1
```{r m15, results='asis'}
dns <- dd %>%
  filter(type_of_lesion == 'No Stroke')

m15 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m15)
```

* Model 2
```{r m16, results='asis'}
m16 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m16)
```

* Model 3
```{r m17, results='asis'}
m17 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m17)
```

* Model 4
```{r m18, results='asis'}
m18 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m18)
```
\newpage

### TMAO Continuous

* Model 1
```{r m150, results='asis'}
m150 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m150)
```

* Model 2
```{r m160, results='asis'}
m160 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m160)
```

* Model 3
```{r m170, results='asis'}
m170 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m170)
```

* Model 4
```{r m180, results='asis'}
m180 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes,
          random = ~1 | center, data=dns,
          na.action=na.exclude)

print_lme_fixed(m180)
```

\newpage

# Subgroup analyses to explore attrition bias

## Follow-up year 3 analysis set (FUP3) (n=2070)

Exclusion of all patients lost to FU before FU-3, had less than 2 cognitive testing sets before FU-3  

### TMAO Quintiles

* Model 1
```{r m21, results='asis'}
fup3 <- FUP3 %>%
  filter(visit.name == 'Baseline')

m21 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=fup3,
          na.action=na.exclude)
print_lme_fixed(m21)
```

* Model 2
```{r m22, results='asis'}
m22 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m22)
```

* Model 3
```{r m23, results='asis'}
m23 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease +prev.stroke.tia,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m23)
```

* Model 4
```{r m24, results='asis'}
m24 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m24)
```
\newpage

### TMAO Continuous

* Model 1
```{r m210, results='asis'}
m210 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m210)
```

* Model 2
```{r m220, results='asis'}
m220 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m220)
```

* Model 3
```{r m230, results='asis'}
m230 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m230)
```

* Model 4
```{r m240, results='asis'}
m240 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m240)
```

\newpage

## Worst analysis set (n=146)  
FAS-FUP3 = all patients, who died before FU-3 and had fewer less than 2 cognitive testing sets before FU-3  

### TMAO Quintiles

* Model 1
```{r m25, results='asis'}
wgs <- WGS %>%
  filter(visit.name == 'Baseline')

m25 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m25)
```

* Model 2
```{r m26, results='asis'}
m26 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m26)
```

* Model 3
```{r m27, results='asis'}
m27 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m27)
```

* Model 4
```{r m28, results='asis'}
m28 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m28)
```
\newpage

### TMAO Continuous

* Model 1
```{r m250, results='asis'}
m250 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m250)
```

* Model 2
```{r m260, results='asis'}
m260 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m260)
```

* Model 3
```{r m270, results='asis'}
m270 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m270)
```

* Model 4
```{r m280, results='asis'}
m280 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m280)
```

\newpage

