---
title: "Association between Trimethylamine N-Oxide (TMAO) and longitudinal cognitive function  in Patients with Atrial Fibrillation - Part 2: Survival and sensitivity analyses"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
  html_document:
    toc: true
    toc_depth: '4'
    df_print: paged
header-includes: \usepackage[some]{background}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, error=F, message=F, warning=F, cache=T)

library(dplyr)
library(readr)
library(nlme)
library(knitr)
library(survival)
options(scipen=999)


# Fixed-effect table for nlme::lme models.
# Uses nlme directly and deliberately does not print sample size.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    tt$`p-value` < 0.001,
    "<0.001",
    formatC(tt$`p-value`, format = "f", digits = digits)
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "r", "c", "r", "r", "r")
    )
  )
}

# Numeric table for survival::coxph models.
# Uses model coefficients directly and deliberately does not print N or event counts.
print_cox_table <- function(model, digits = 3) {
  beta <- stats::coef(model)
  vc <- stats::vcov(model)

  # Keep only coefficients with matching variance estimates.
  common <- intersect(names(beta), rownames(vc))
  beta <- beta[common]
  se <- sqrt(diag(vc[common, common, drop = FALSE]))

  z <- beta / se
  p <- 2 * stats::pnorm(abs(z), lower.tail = FALSE)
  lower <- exp(beta - stats::qnorm(0.975) * se)
  upper <- exp(beta + stats::qnorm(0.975) * se)

  fmt_num <- function(x) {
    ifelse(is.na(x), NA_character_, formatC(x, format = "f", digits = digits))
  }
  fmt_p <- function(x) {
    ifelse(
      is.na(x),
      NA_character_,
      ifelse(x < 0.001, "<0.001", formatC(x, format = "f", digits = digits))
    )
  }

  out <- data.frame(
    Parameter = names(beta),
    HR = fmt_num(exp(beta)),
    `95% CI` = paste0(fmt_num(lower), " to ", fmt_num(upper)),
    `p value` = fmt_p(p),
    check.names = FALSE,
    row.names = NULL
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "c", "r")
    )
  )
}

```

```{r Data Wrangling}
dd <- read_csv("Fu5.csv")
dd$tmao_log <- log(dd$tmao + 1)
dd$tmao_q <- as.factor(dd$tmao_q)

dd <- dd %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

dd <- dd %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)

FUP3 <- read_csv("FUP3.long.csv")
FUP3 <- FUP3 %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
FUP3$tmao_log <- log(FUP3$tmao + 1)
FUP3$tmao_q <- as.factor(FUP3$tmao_q)

FUP3 <- FUP3 %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

FUP3 <- FUP3 %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)

WGS <- read_csv("WGS.long.csv")
WGS <- WGS %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
WGS$tmao_log <- log(WGS$tmao + 1)
WGS$tmao_q <- as.factor(WGS$tmao_q)

WGS <- WGS %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

WGS <- WGS %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)

WGS$time_to_CoCo_change <- as.numeric(WGS$time_to_CoCo_change)

```

\newpage

# Association between TMAO and time to cognitive decline (Survival analyses)

## Mild Cognitive Impairment (MoCA < 26)

### TMAO Quintiles

```{r c1234,results="asis"}

c4a <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c4a)

c4b <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c4b)

```


\newpage

### TMAO continuous

```{r c12340,results="asis"}

c4a0 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c4a0)

c4b0 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c4b0)

```

\newpage

## Early dementia (MoCA < 21)

### TMAO Quintiles

```{r c5678,results="asis"}

c8a <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c8a)

c8b <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c8b)

```

\newpage

### TMAO continuous

```{r c56780,results="asis"}

c80a <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c80a)

c80b <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c80b)

```

\newpage

## CoCo score change >= 1 SD -> event of cognitive decline

### TMAO Quintiles

```{r c1,results="asis"}

c14a <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c14a)

c14b <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c14b)

```


\newpage

### TMAO continuous

```{r c10,results="asis"}

c140a <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c140a)

c140b <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c140b)

```

\newpage

## TMT-A change >= 1 SD

### TMAO quintiles

```{r c2,results="asis"}

c18a <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c18a)

c18b <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c18b)

```

\newpage

### TMAO Continuous

```{r c20,results="asis"}

c180a <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c180a)


c180b <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c180b)


```

\newpage

## TMT-B change >= 1 SD

### TMAO Quintiles 
```{r c3,results="asis"}

c24a <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c24a)

c24b <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c24b)

```


\newpage 

### TMAO Continuous

```{r c30,results="asis"}

c240a <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_log
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c240a)

c240b <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_log
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c240b)

```

\newpage

## SF score change >= 1 SD

### TMAO Quintiles

```{r c4,results="asis"}

c28a <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c28a)

c28b <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c28b)

```


\newpage

### TMAO continuous

```{r c40,results="asis"}

c280a <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_log  
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c280a)

c280b <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_log  
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = dd) 

print_cox_table(c280b)
```

\newpage

## DSST change >= 1 SD

### TMAO Quintiles

```{r c5,results="asis"}

c34a <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
           + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c34a)

c34b <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c34b)

```

\newpage

### TMAO Continuous

```{r c50,results="asis"}

c340a <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c340a)

c340b <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c340b)

```
\newpage

## MoCA change >=1 SD

### TMAO Quintiles

```{r c6,results="asis"}

c38a <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c38a)

c38b <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c38b)

```

\newpage

### TMAO continuous
```{r c60,results="asis"}

c380a <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c380a)

c380b <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = dd) 

print_cox_table(c380b)

```

\newpage

#Sensitivity Analysis Short-term and long-term group

## Long-term group (n=2070)

**Linear mixed effect model**

### CoCo overall

#### TMAO Quintiles

* Model 4a
```{r m34a, results='asis'}
m34a <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m34a)
```


* Model 4b
```{r m34b, results='asis'}
m34b <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m34b)
```

\newpage

#### TMAO Continuous

* Model 4a
```{r m340a, results='asis'}
m340a <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m340a)
```


* Model 4b
```{r m340b, results='asis'}
m340b <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m340b)
```

\newpage

### MoCA score

#### TMAO Quintiles

* Model 4a
```{r m38a, results='asis'}
m38a <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m38a)
```


* Model 4b
```{r m38b, results='asis'}
m38b <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m38b)
```

\newpage

#### TMAO Continuous

* Model 4a
```{r m38a0, results='asis'}
m38a0 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m38a0)
```

* Model 4b
```{r m38b0, results='asis'}
m38b0 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m38b0)
```

\newpage

**Time-to-event/ Survival analysis**

### Mild Cognitive Impairment (MoCA < 26)

#### TMAO Quintiles

```{r c7,results="asis"}

c44a <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c44a)

c44b <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = FUP3) 

print_cox_table(c44b)

```
\newpage

#### TMAO continuous

```{r c70,results="asis"}

c440a <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c440a)

c440b <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = FUP3) 

print_cox_table(c440b)
```

\newpage

### CoCo score change >= 1 SD 

#### TMAO Quintiles

```{r c8,results="asis"}

c48a <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c48a)

c48b <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = FUP3) 

print_cox_table(c48b)

```
\newpage

### TMAO continuous

```{r c80,results="asis"}

c480a <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c480a)

c480b <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = FUP3) 

print_cox_table(c480b)

```
\newpage

## Short-term group (n=146)  

The sample size is too small for CoCo change >= 1 SD and also too small for overall cognition analyses in linear mixed effect model  

**Time-to-event/ Survival analysis**

### Mild Cognitive Impairment (MoCA < 26)

#### TMAO Quintiles

```{r c9,results="asis"}

c54a <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = WGS) 

print_cox_table(c54a)

c54b <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes 
            + frailty(center), data = WGS) 

print_cox_table(c54b)

```
\newpage

#### TMAO continuous

```{r c90,results="asis"}

c540a <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + creatinine.gfr
            + frailty(center), data = WGS) 

print_cox_table(c540a)

c540b <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes
            + frailty(center), data = WGS) 

print_cox_table(c540b)
```

\newpage
