---
title: "Association between Trimethylamine N-Oxide (TMAO) and longitudinal cognitive function  in Patients with Atrial Fibrillation"
author: "P. Lee"
date: "11.03.2025"
output:
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: false
  html_document:
    toc: true
    toc_depth: 4
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE
)

library(tidyverse)
library(nlme)
library(survival)
library(survminer)
library(knitr)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
# Uses nlme directly to avoid version-sensitive extraction of random-effect
# variances by parameters/datawizard. Sample size is deliberately not printed.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "r", "c", "r", "r", "r")
    )
  )
}

# Numeric fixed-effect table for survival::coxph models.
# Avoids forestmodel/ggplot compatibility issues and deliberately does not
# print sample size or event counts.
print_cox_table <- function(model, digits = 3) {
  beta <- stats::coef(model)
  vc <- stats::vcov(model)

  common <- intersect(names(beta), rownames(vc))
  beta <- beta[common]
  se <- sqrt(diag(vc[common, common, drop = FALSE]))

  z <- beta / se
  p <- 2 * stats::pnorm(abs(z), lower.tail = FALSE)
  lower <- exp(beta - stats::qnorm(0.975) * se)
  upper <- exp(beta + stats::qnorm(0.975) * se)

  fmt_num <- function(x) {
    ifelse(
      is.na(x),
      NA_character_,
      formatC(x, format = "f", digits = digits)
    )
  }

  fmt_p <- function(x) {
    ifelse(
      is.na(x),
      NA_character_,
      ifelse(
        x < 0.001,
        "<0.001",
        formatC(x, format = "f", digits = digits)
      )
    )
  }

  out <- data.frame(
    Parameter = names(beta),
    HR = fmt_num(exp(beta)),
    `95% CI` = paste0(fmt_num(lower), " to ", fmt_num(upper)),
    `p value` = fmt_p(p),
    check.names = FALSE,
    row.names = NULL
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "c", "r")
    )
  )
}
```

```{r data-wrangling}
dd <- read_csv("Fu5.csv")
dd$tmao_log <- log1p(dd$tmao )
dd$tmao_q <- as.factor(dd$tmao_q)

dd <- dd %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

dd <- dd %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)

FUP3 <- read_csv("FUP3.long.csv")
FUP3 <- FUP3 %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
FUP3$tmao_log <- log1p(FUP3$tmao )
FUP3$tmao_q <- as.factor(FUP3$tmao_q)

FUP3 <- FUP3 %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

FUP3 <- FUP3 %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)

WGS <- read_csv("WGS.long.csv")
WGS <- WGS %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
WGS$tmao_log <- log1p(WGS$tmao )
WGS$tmao_q <- as.factor(WGS$tmao_q)

WGS <- WGS %>%
  mutate(pat.id = as.factor(pat.id),
         visit.name = as.factor(visit.name),
         pat.sex = as.factor(pat.sex),
         vhf.typ.aktuell.bl = as.factor(vhf.typ.aktuell.bl),
         rauchen = as.factor(rauchen),
         stop.grund = as.factor(stop.grund),
         prev.hypertonie = as.factor(prev.hypertonie),
         prev.diabetes = as.factor(prev.diabetes),
         prev.stroke.tia = as.factor(prev.stroke.tia),
         prev.coronary.heart.disease = as.factor(prev.coronary.heart.disease),
         sport = as.factor(sport),
         highest.education.level.groups = as.factor(highest.education.level.groups),
         type_of_lesion = as.factor(type_of_lesion),
         dementia = as.factor(dementia),
         cog_impairment = as.factor(cog_impairment),
         MoCA_change = as.factor(MoCA_change),
         CoCo_change = as.factor(CoCo_change),
         DSST_change = as.factor(DSST_change),
         SF_change = as.factor(SF_change),
         TMTA_change = as.factor(TMTA_change),
         TMTB_change = as.factor(TMTB_change))

WGS <- WGS %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)

WGS$time_to_CoCo_change <- as.numeric(WGS$time_to_CoCo_change)

```

\newpage

# Association between TMAO and overall cognitive function over time (LME)

**linear mixed effects model (LME)** is used according to manual page 6 (structure model I)  

* Fixed Effects

Intercept: This is the predicted value of Cognitive score when both time.since.first.measurement and TMAO are zero.  
      
Time.since.first.measurement: This coefficient represents the estimated change in the cognitive score per unit increase in time (e.g., per year) when TMAO is at its reference level (typically zero or its centered value).  
Example: If the coefficient is –0.5, then for a patient with average (or reference) TMAO, the cognitive score declines by 0.5 points per time unit.  

TMAO: This term indicates how a one-unit change in TMAO is associated with the cognitive score at baseline (time = 0).  
Example: If the coefficient is –0.2, it suggests that a one-unit higher TMAO is associated with a 0.2-point lower cognitive score at baseline.  

Interaction (time.since.first.measurement:TMAO): This term shows whether the effect of time on cognitive score depends on the level of TMAO.  
Example: Suppose the coefficient is –0.05. For every one-unit increase in TMAO, the decline in cognitive score per time unit is 0.05 points greater.  
Putting it Together: For a patient with TMAO value xx, the estimated slope for time becomes:  
slope=(time coefficient)+x×(interaction coefficient)  
If the time coefficient is –0.5 and TMAO is 3, then the slope would be –0.5 + 3 × (–0.05) = –0.65, meaning that higher TMAO is associated with a faster decline.  

* Random Effects

The model includes random effects to account for individual variability:  
The term (time.since.first.measurement | center/pat.id) means that each patient (nested within centers) is allowed to have their own baseline cognitive score (intercept) and their own rate of change over time (slope for time).  
This helps account for differences between patients. For example, some patients might naturally have higher cognitive scores or decline at a different rate than the average estimated by the fixed effects.  

* Covariates
Model 1 (epidemiologic variables): age, sex, education (years), quality of life (EQ-ED-5L), geriatric depression scale (GDS)  
Model 2: (modifiable cardiovascular risk factors): Model 1 + smoking status, BMI, hypertension, alcohol consumption, physical activity  
Model 3 (cardiovascular events): Model 1 + Model 2 + coronary artery disease, prev. stroke   
Model 4 (metabolic variables): Model 1 + Model 2 + Model 3 + diabetes status, renal function (eGFR) 
\newpage

## CoCo score

### TMAO quintiles

* Model 1
```{r m1, results='asis'}
m1 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m1)
```

* Model 2
```{r m2, results='asis'}
m2 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m2)
```

* Model 3
```{r m3, results='asis'}
m3 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m3)
```

* Model 4
```{r m4, results='asis'}
m4 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m4)
```
\newpage

### TMAO Continuous

* Model 1
```{r m10, results='asis'}
m10 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m10)
```

* Model 2
```{r m20, results='asis'}
m20 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m20)
```

* Model 3
```{r m30, results='asis'}
m30 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m30)
```

* Model 4
```{r m40, results='asis'}
m40 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m40)
```

\newpage

## MoCA score

### TMAO Quintiles

* Model 1
```{r m5, results='asis'}
m5 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m5)
```

* Model 2
```{r m6, results='asis'}
m6 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m6)
```

* Model 3
```{r m7, results='asis'}
m7 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m7)
```

* Model 4
```{r m8, results='asis'}
m8 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m8)
```


\newpage

### TMAO Continuous

* Model 1
```{r m50, results='asis'}
m50 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m50)
```

* Model 2
```{r m60, results='asis'}
m60 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m60)
```

* Model 3
```{r m70, results='asis'}
m70 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m70)
```

* Model 4
```{r m80, results='asis'}
m80 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m80)
```

\newpage

## TMT-A score

### TMAO Quintiles 

* Model 1
```{r m11, results='asis'}
m11 <- lme(TMTA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m11)
```

* Model 2
```{r m12, results='asis'}
m12 <- lme(TMTA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m12)
```

* Model 3
```{r m13, results='asis'}
m13 <- lme(TMTA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m13)
```

* Model 4
```{r m14, results='asis'}
m14 <- lme(TMTA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m14)
```


\newpage

### TMAO Continuous

* Model 1
```{r m110, results='asis'}
m110 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m110)
```

* Model 2
```{r m120, results='asis'}
m120 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m120)
```

* Model 3
```{r m130, results='asis'}
m130 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m130)
```

* Model 4
```{r m140, results='asis'}
m140 <- lme(TMTA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m140)
```

\newpage

## TMT-B score

### TMAO Quintiles

* Model 1
```{r m15, results='asis'}
m15 <- lme(TMTB ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m15)
```

* Model 2
```{r m16, results='asis'}
m16 <- lme(TMTB ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m16)
```

* Model 3
```{r m17, results='asis'}
m17 <- lme(TMTB ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m17)
```

* Model 4
```{r m18, results='asis'}
m18 <- lme(TMTB ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m18)
```


\newpage

### TMAO Continuous

* Model 1
```{r m150, results='asis'}
m150 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m150)
```

* Model 2
```{r m160, results='asis'}
m160 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m160)
```

* Model 3
```{r m170, results='asis'}
m170 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m170)
```

* Model 4
```{r m180, results='asis'}
m180 <- lme(TMTB ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m180)
```


\newpage

## SF score

### TMAO Quintiles

* Model 1
```{r m21, results='asis'}
m21 <- lme(SF ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m21)
```

* Model 2
```{r m22, results='asis'}
m22 <- lme(SF ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m22)
```

* Model 3
```{r m23, results='asis'}
m23 <- lme(SF ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m23)
```

* Model 4
```{r m24, results='asis'}
m24 <- lme(SF ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m24)
```


\newpage

### TMAO Continuous

* Model 1
```{r m210, results='asis'}
m210 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m210)
```

* Model 2
```{r m220, results='asis'}
m220 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m220)
```

* Model 3
```{r m230, results='asis'}
m230 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m230)
```

* Model 4
```{r m240, results='asis'}
m240 <- lme(SF ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m240)
```

\newpage 

## DSST score

### TMAO Quintiles

* Model 1
```{r m25, results='asis'}
m25 <- lme(DSST ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m25)
```

* Model 2
```{r m26, results='asis'}
m26 <- lme(DSST ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m26)
```

* Model 3
```{r m27, results='asis'}
m27 <- lme(DSST ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m27)
```

* Model 4
```{r m28, results='asis'}
m28 <- lme(DSST ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m28)
```


\newpage

### TMAO Continuous

* Model 1
```{r m250, results='asis'}
m250 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m250)
```

* Model 2
```{r m260, results='asis'}
m260 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m260)
```

* Model 3
```{r m270, results='asis'}
m270 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m270)
```

* Model 4
```{r m280, results='asis'}
m280 <- lme(DSST ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=dd,
          na.action=na.exclude)

print_lme_fixed(m280)
```

# Association between TMAO and time to cognitive decline (Survival analyses)

## Mild Cognitive Impairment (MoCA < 26)

### TMAO Quintiles

```{r c1234,results="asis", fig.width=12, fig.height=8}

c1 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c1)

c2 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c2)

c3 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c3)

c4 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c4)

```

**Figure: Kaplan-Meier Free from cognitive impairment survival curve**

```{r km-plot-1,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(cogimpair.y = time_to_cog_impairment/365.25)

km1 <- survfit(Surv(cogimpair.y, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q, data = dd)
ggsurvplot(km1, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```

\newpage

### TMAO continuous

```{r c12340,results="asis", fig.width=12, fig.height=8}

c10 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c10)

c20 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c20)

c30 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c30)

c40 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c40)

```

\newpage

## Early dementia (MoCA < 21)

### TMAO Quintiles

```{r c5678,results="asis", fig.width=12, fig.height=8}

c5 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c5)

c6 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c6)

c7 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c7)

c8 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c8)

```

**Figure: Kaplan-Meier Free from early dementia survival curve**

```{r km-plot-2,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(dementia.y = time_to_dementia/365.25)

km2 <- survfit(Surv(dementia.y, 
                 as.numeric(as.character(dementia))) ~ tmao_q, data = dd)
ggsurvplot(km2, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```

\newpage

### TMAO continuous

```{r c56780,results="asis", fig.width=12, fig.height=8}

c50 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c50)

c60 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c60)

c70 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c70)

c80 <- coxph(Surv(time_to_dementia, 
                 as.numeric(as.character(dementia))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c80)

```

\newpage

## CoCo score change >= 1 SD -> event of cognitive decline

### TMAO Quintiles

```{r c1,results="asis", fig.width=12, fig.height=8}

c11 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c11)

c12 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c12)

c13 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c13)

c14 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c14)

```

**Figure: Kaplan-Meier Free from CoCo change survival curve**

```{r km-plot-3,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(CoCo_change.y = time_to_CoCo_change/365.25)

km3 <- survfit(Surv(CoCo_change.y, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q, data = dd)
ggsurvplot(km3, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```

\newpage

### TMAO continuous

```{r c10,results="asis", fig.width=12, fig.height=8}

c110 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c110)

c120 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c120)

c130 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c130)

c140 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c140)

```

\newpage

## TMT-A change >= 1 SD

### TMAO quintiles

```{r c2,results="asis", fig.width=12, fig.height=8}

c15 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c15)

c16 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c16)

c17 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c17)

c18 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c18)

```

```{r km-plot-4,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(TMTA_change.y = time_to_TMTA_change/365.25)

km4 <- survfit(Surv(TMTA_change.y, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_q, data = dd)
ggsurvplot(km4, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```
\newpage

### TMAO Continuous

```{r c20,results="asis", fig.width=12, fig.height=8}

c150 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c150)

c160 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_log  
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c160)

c170 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_log  
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c170)

c180 <- coxph(Surv(time_to_TMTA_change, 
                 as.numeric(as.character(TMTA_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c180)

```

\newpage

## TMT-B change >= 1 SD

### TMAO Quintiles 
```{r c3,results="asis", fig.width=12, fig.height=8}

c21 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c21)

c22 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c22)

c23 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c23)

c24 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c24)

```

```{r km-plot-5,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(TMTB_change.y = time_to_TMTB_change/365.25)

km5 <- survfit(Surv(TMTB_change.y, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_q, data = dd)
ggsurvplot(km5, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```

\newpage 

### TMAO Continuous

```{r c30,results="asis", fig.width=12, fig.height=8}

c210 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c210)

c220 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_log
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c220)

c230 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c230)

c240 <- coxph(Surv(time_to_TMTB_change, 
                 as.numeric(as.character(TMTB_change))) ~ tmao_log
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c240)

```

\newpage

## SF score change >= 1 SD

### TMAO Quintiles

```{r c4,results="asis", fig.width=12, fig.height=8}

c25 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c25)

c26 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c26)

c27 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c27)

c28 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c28)

```

```{r km-plot-6,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(SF_change.y = time_to_SF_change/365.25)

km6 <- survfit(Surv(SF_change.y, 
                 as.numeric(as.character(SF_change))) ~ tmao_q, data = dd)
ggsurvplot(km6, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```

\newpage

### TMAO continuous

```{r c40,results="asis", fig.width=12, fig.height=8}

c250 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c250)

c260 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_log  
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c260)

c270 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c270)

c280 <- coxph(Surv(time_to_SF_change, 
                 as.numeric(as.character(SF_change))) ~ tmao_log  
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c280)
```

\newpage

## DSST change >= 1 SD

### TMAO Quintiles

```{r c5,results="asis", fig.width=12, fig.height=8}

c31 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c31)

c32 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c32)

c33 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c33)

c34 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c34)

```

```{r km-plot-7,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(DSST_change.y = time_to_DSST_change/365.25)

km7 <- survfit(Surv(DSST_change.y, 
                 as.numeric(as.character(DSST_change))) ~ tmao_q, data = dd)
ggsurvplot(km7, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```
\newpage

### TMAO Continuous

```{r c50,results="asis", fig.width=12, fig.height=8}

c310 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c310)

c320 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c320)

c330 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c330)

c340 <- coxph(Surv(time_to_DSST_change , 
                 as.numeric(as.character(DSST_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c340)

```
\newpage

## MoCA change >=1 SD

### TMAO Quintiles

```{r c6,results="asis", fig.width=12, fig.height=8}

c35 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c35)

c36 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c36)

c37 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c37)

c38 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c38)

```

```{r km-plot-8,results="asis", fig.show='asis', fig.height=5, fig.width=7}
dd <- dd %>%
  mutate(MoCA_change.y = time_to_MoCA_change/365.25)

km8 <- survfit(Surv(MoCA_change.y, 
                 as.numeric(as.character(MoCA_change))) ~ tmao_q, data = dd)
ggsurvplot(km8, data=dd, 
           break.x.by = 1, 
           legend.title = "TMAO Quintiles",
           legend.labs = c("Q1", "Q2", "Q3", "Q4", 'Q5'),
           xlab = "Time (years)",
           ylab = "Probability of Cognitive impairment-Free survival",
           pval = TRUE,
           conf.int = TRUE,
           palette = c("lightcoral", "tomato1", "tomato3", "tomato4", 'red4'),
           ggtheme = theme_classic())
```

\newpage

### TMAO continuous
```{r c60,results="asis", fig.width=12, fig.height=8}

c350 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = dd)

print_cox_table(c350)

c360 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = dd) 

print_cox_table(c360)

c370 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = dd) 

print_cox_table(c370)

c380 <- coxph(Surv(time_to_MoCA_change , 
                 as.numeric(as.character(MoCA_change ))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = dd) 

print_cox_table(c380)
```

\newpage

#Subgroup analyses to explore attrition bias

## Follow-up year 3 analysis set (FUP3)
Exclusion of all patients lost to FU before FU-3, had less than 2 cognitive testing sets before FU-3  

**Linear mixed effect model**

### CoCo overall

#### TMAO Quintiles

* Model 1
```{r m31, results='asis'}
m31 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m31)
```

* Model 2
```{r m32, results='asis'}
m32 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m32)
```

* Model 3
```{r m33, results='asis'}
m33 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m33)
```

* Model 4
```{r m34, results='asis'}
m34 <- lme(CoCo ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m34)
```

\newpage

#### TMAO Continuous

* Model 1
```{r m310, results='asis'}
m310 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m310)
```

* Model 2
```{r m320, results='asis'}
m320 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m320)
```

* Model 3
```{r m330, results='asis'}
m330 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m330)
```

* Model 4
```{r m340, results='asis'}
m340 <- lme(CoCo ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m340)
```

\newpage

### MoCA score

#### TMAO Quintiles

* Model 1
```{r m35, results='asis'}
m35 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m35)
```

* Model 2
```{r m36, results='asis'}
m36 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m36)
```

* Model 3
```{r m37, results='asis'}
m37 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m37)
```

* Model 4
```{r m38, results='asis'}
m38 <- lme(MoCA ~ time_since_first + tmao_q +
                 time_since_first:tmao_q +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m38)
```

\newpage

#### TMAO Continuous

* Model 1
```{r m350, results='asis'}
m350 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m350)
```

* Model 2
```{r m360, results='asis'}
m360 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m360)
```

* Model 3
```{r m370, results='asis'}
m370 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m370)
```

* Model 4
```{r m380, results='asis'}
m380 <- lme(MoCA ~ time_since_first + tmao_log +
                 time_since_first:tmao_log +
                 age.bl + pat.sex + highest.education.level.groups + gesundheitszustand +
                 gds.total + 
                 rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
                 + prev.coronary.heart.disease + prev.stroke.tia
                 + prev.diabetes + creatinine.gfr,
          random = ~ time_since_first | center/pat.id, data=FUP3,
          na.action=na.exclude)

print_lme_fixed(m380)
```

\newpage

**Time-to-event/ Survival analysis**

### Mild Cognitive Impairment (MoCA < 26)

#### TMAO Quintiles

```{r c7,results="asis", fig.width=12, fig.height=8}

c41 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = FUP3)

print_cox_table(c41)

c42 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = FUP3) 

print_cox_table(c42)

c43 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = FUP3) 

print_cox_table(c43)

c44 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c44)

```
\newpage

#### TMAO continuous

```{r c70,results="asis", fig.width=12, fig.height=8}

c410 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = FUP3)

print_cox_table(c410)

c420 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = FUP3) 

print_cox_table(c420)

c430 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = FUP3) 

print_cox_table(c430)

c440 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c440)
```

\newpage

### CoCo score change >= 1 SD 

#### TMAO Quintiles

```{r c8,results="asis", fig.width=12, fig.height=8}

c45 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = FUP3)

print_cox_table(c45)

c46 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = FUP3) 

print_cox_table(c46)

c47 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = FUP3) 

print_cox_table(c47)

c48 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c48)

```
\newpage

### TMAO continuous

```{r c80,results="asis", fig.width=12, fig.height=8}

c450 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = FUP3)

print_cox_table(c450)

c460 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = FUP3) 

print_cox_table(c460)

c470 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = FUP3) 

print_cox_table(c470)

c480 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = FUP3) 

print_cox_table(c480)

```
\newpage

## Worst analysis set  
FAS-FUP3 = all patients, who died before FU-3 and had fewer less than 2 cognitive testing sets before FU-3  

The sample size is too small for CoCo change >= 1 SD and also too small for overall cognition analyses in linear mixed effect model  

**Time-to-event/ Survival analysis**

### Mild Cognitive Impairment (MoCA < 26)

#### TMAO Quintiles

```{r c9,results="asis", fig.width=12, fig.height=8}

c51 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = WGS)

print_cox_table(c51)

c52 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = WGS) 

print_cox_table(c52)

c53 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = WGS) 

print_cox_table(c53)

c54 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = WGS) 

print_cox_table(c54)

```
\newpage

#### TMAO continuous

```{r c90,results="asis", fig.width=12, fig.height=8}

c510 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = WGS)

print_cox_table(c510)

c520 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = WGS) 

print_cox_table(c520)

c530 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = WGS) 

print_cox_table(c530)

c540 <- coxph(Surv(time_to_cog_impairment, 
                 as.numeric(as.character(cog_impairment))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = WGS) 

print_cox_table(c540)
```

\newpage

