---
title: "Updated analyses"
author: "P. Lee"
date: "27.10.2025"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
    number_sections: no
  word_document:
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE
)

library(tidyverse)
library(nlme)
library(survival)
library(knitr)

options(scipen = 999)

# Fixed-effect table for nlme::lme models.
# Uses nlme directly to avoid version-sensitive extraction of random-effect
# variances by parameters/datawizard. Sample size is deliberately not printed.
print_lme_fixed <- function(model, digits = 3) {
  tt <- as.data.frame(summary(model)$tTable)
  tt$Parameter <- rownames(tt)
  rownames(tt) <- NULL

  crit <- qt(0.975, df = tt$DF)
  tt$CI_low <- tt$Value - crit * tt$Std.Error
  tt$CI_high <- tt$Value + crit * tt$Std.Error

  p_fmt <- ifelse(
    is.na(tt$`p-value`),
    NA_character_,
    ifelse(
      tt$`p-value` < 0.001,
      "<0.001",
      formatC(tt$`p-value`, format = "f", digits = digits)
    )
  )

  out <- data.frame(
    Parameter = tt$Parameter,
    Estimate = round(tt$Value, digits),
    SE = round(tt$Std.Error, digits),
    `95% CI` = sprintf(
      paste0("%.", digits, "f to %.", digits, "f"),
      tt$CI_low, tt$CI_high
    ),
    DF = tt$DF,
    `t value` = round(tt$`t-value`, digits),
    `p value` = p_fmt,
    check.names = FALSE
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "r", "c", "r", "r", "r")
    )
  )
}

# Numeric fixed-effect table for survival::coxph models.
# Avoids forestmodel/ggplot compatibility issues and deliberately does not
# print sample size or event counts.
print_cox_table <- function(model, digits = 3) {
  beta <- stats::coef(model)
  vc <- stats::vcov(model)

  common <- intersect(names(beta), rownames(vc))
  beta <- beta[common]
  se <- sqrt(diag(vc[common, common, drop = FALSE]))

  z <- beta / se
  p <- 2 * stats::pnorm(abs(z), lower.tail = FALSE)
  lower <- exp(beta - stats::qnorm(0.975) * se)
  upper <- exp(beta + stats::qnorm(0.975) * se)

  fmt_num <- function(x) {
    ifelse(
      is.na(x),
      NA_character_,
      formatC(x, format = "f", digits = digits)
    )
  }

  fmt_p <- function(x) {
    ifelse(
      is.na(x),
      NA_character_,
      ifelse(
        x < 0.001,
        "<0.001",
        formatC(x, format = "f", digits = digits)
      )
    )
  }

  out <- data.frame(
    Parameter = names(beta),
    HR = fmt_num(exp(beta)),
    `95% CI` = paste0(fmt_num(lower), " to ", fmt_num(upper)),
    `p value` = fmt_p(p),
    check.names = FALSE,
    row.names = NULL
  )

  print(
    knitr::kable(
      out,
      align = c("l", "r", "c", "r")
    )
  )
}
```

```{r data-wrangling}
dd <- read_csv("Baseline.csv")
dd <- dd %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
dd$tmao_log <- log1p(dd$tmao )
dd$tmao_q <- as.factor(dd$tmao_q)

ddns <- dd %>%
  filter(type_of_lesion == 'No Stroke')
ddos <- dd %>%
  filter(type_of_lesion == 'Overt Stroke')
ddss <- dd %>%
  filter(type_of_lesion == 'Silent Stroke')

FUP3 <- read_csv("FUP3.csv")
FUP3 <- FUP3 %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
FUP3$tmao_log <- log1p(FUP3$tmao )
FUP3$tmao_q <- as.factor(FUP3$tmao_q)

FUP3 <- FUP3 %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)

WGS <- read_csv("WGS.csv")
WGS <- WGS %>%
  mutate(tmao_q = as.factor(ntile(tmao,5)))
WGS$tmao_log <- log1p(WGS$tmao )
WGS$tmao_q <- as.factor(WGS$tmao_q)

WGS <- WGS %>%
  group_by(pat.id) %>%
  # Extract the baseline creatinine.gfr value for each patient
  mutate(baseline_creatinine = creatinine.gfr[visit.name == "Baseline"][1]) %>%
  # Replace missing creatinine.gfr with the baseline value
  mutate(creatinine.gfr = ifelse(is.na(creatinine.gfr), baseline_creatinine, creatinine.gfr)) %>%
  ungroup() %>%
  # Optionally, remove the temporary column
  select(-baseline_creatinine)
```

# Sensitivity analyses for TMAO with HV for LTG and STG  

Model 1: age, sex, total intracranial volume, education level  
Model 2: model 1 + smoking, bmi, HT, physical activity + alcohol  
Model 3: model 2 + History of CAD + History of stroke/TIA  
Model 4: model 3 + DM + eGFR  

## Long-term group  
### TMAO Quintiles  

* Model 1
```{r m21, results='asis'}
fup3 <- FUP3 %>%
  filter(visit.name == 'Baseline')

m21 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m21)
```

* Model 2
```{r m22, results='asis'}
m22 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m22)
```

* Model 3
```{r m23, results='asis'}
m23 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m23)
```

* Model 4
```{r m24, results='asis'}
m24 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m24)
```
\newpage

### TMAO Continuous  

* Model 1
```{r m210, results='asis'}
m210 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m210)
```

* Model 2
```{r m220, results='asis'}
m220 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m220)
```

* Model 3
```{r m230, results='asis'}
m230 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m230)
```

* Model 4
```{r m240, results='asis'}
m240 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=fup3,
          na.action=na.exclude)

print_lme_fixed(m240)
```

\newpage

## Short term group  

### TMAO Quintiles

* Model 1
```{r m25, results='asis'}
wgs <- WGS %>%
  filter(visit.name == 'Baseline')

m25 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m25)
```

* Model 2
```{r m26, results='asis'}
m26 <- lme(HVraw ~ tmao_q  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m26)
```

* Model 3
```{r m27, results='asis'}
m27 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m27)
```

* Model 4
```{r m28, results='asis'}
m28 <- lme(HVraw ~ tmao_q +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m28)
```
\newpage

### TMAO Continuous

* Model 1
```{r m250, results='asis'}
m250 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m250)
```

* Model 2
```{r m260, results='asis'}
m260 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m260)
```

* Model 3
```{r m270, results='asis'}
m270 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m270)
```

* Model 4
```{r m280, results='asis'}
m280 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease + prev.stroke.tia
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=wgs,
          na.action=na.exclude)

print_lme_fixed(m280)
```

\newpage

# Sensitivity analyses for TMAO with CoCo score change >= 1 SD

## Long-term group

### TMAO Quintiles

```{r c1,results="asis", fig.width=12, fig.height=8}
c11 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = FUP3)

c12 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = FUP3) 

c13 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = FUP3) 


c14 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_q 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = FUP3) 


# print table(s)
print_cox_table(c11)
print_cox_table(c12)
print_cox_table(c13)
print_cox_table(c14)

```

\newpage

### TMAO continuous

```{r c10,results="asis", fig.width=12, fig.height=8}

c110 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total
            + frailty(center), data = FUP3)

c120 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption
             + frailty(center), data = FUP3) 

c130 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
             + frailty(center), data = FUP3) 

c140 <- coxph(Surv(time_to_CoCo_change, 
                 as.numeric(as.character(CoCo_change))) ~ tmao_log 
            + age.bl + pat.sex + highest.education.level.groups 
            + gesundheitszustand + gds.total 
            + rauchen + bmi + prev.hypertonie + sport + total.alcohol.consumption 
            + prev.coronary.heart.disease + prev.stroke.tia
            + prev.diabetes + creatinine.gfr
            + frailty(center), data = FUP3) 

# print table(s)
print_cox_table(c110)
print_cox_table(c120)
print_cox_table(c130)
print_cox_table(c140)

```

\newpage

## Short-term group

No events of CoCo decline >= 1 SD

\newpage

# Association of TMAO with HV (Stroke subgroup analyses)

Model 1 (Epidemiologic variables): Age, Sex, Level of education, Intracranial Volume (sbtiv.vol)    
Model 2 (Modifiable cardiovascular risk factors): Model 1 + smoking, BMI, HT, Alcohol consumption, Physical activity    
Model 3 (Cardiovascular events): Model 1 + Model 2 + Coronary artery disease 
Model 4 (Metabolic variables): Model 1 + Model 2 + Model 3 + Diabetes status + eGFR  

## Overt stroke

### TMAO Continuous

* Model 1
```{r m60, results='asis'}
m60 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m60)
```

* Model 2
```{r m70, results='asis'}
m70 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m70)
```

* Model 3
```{r m80, results='asis'}
m80 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease,
          random = ~1 | center, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m80)
```

* Model 4
```{r m90, results='asis'}
m90 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=ddos,
          na.action=na.exclude)

print_lme_fixed(m90)
```

\newpage

## Silent stroke


### TMAO Continuous

* Model 1
```{r m110, results='asis'}
m110 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m110)
```

* Model 2
```{r m120, results='asis'}
m120 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m120)
```

* Model 3
```{r m130, results='asis'}
m130 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease ,
          random = ~1 | center, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m130)
```

* Model 4
```{r m140, results='asis'}
m140 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=ddss,
          na.action=na.exclude)

print_lme_fixed(m140)
```

\newpage

## No stroke


### TMAO Continuous

* Model 1
```{r m150, results='asis'}
m150 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups,
          random = ~1|center, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m150)
```

* Model 2
```{r m160, results='asis'}
m160 <- lme(HVraw ~ tmao_log  +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption,
          random = ~1 | center, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m160)
```

* Model 3
```{r m170, results='asis'}
m170 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease,
          random = ~1 | center, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m170)
```

* Model 4
```{r m180, results='asis'}
m180 <- lme(HVraw ~ tmao_log +age.bl+pat.sex+sbtiv.vol + highest.education.level.groups
            + rauchen + bmi + prev.hypertonie + sport +
            total.alcohol.consumption + prev.coronary.heart.disease 
          +prev.diabetes + creatinine.gfr,
          random = ~1 | center, data=ddns,
          na.action=na.exclude)

print_lme_fixed(m180)
```

\newpage